---
title: "Oranzees model (02)"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R markdown file documents the development of the Oranzees model (it continues from [01-oranzees.html](01-oranzees.html)). The code-only version to run the model is in the file [main.R](main.R) in the same repository.

## Food-related behaviours

Now that we had implemented the social behaviours, we can work on the food-related ones. As before, we need to define how the state of an oranzee influences its probability to innovate. First, we want that the two nutrients Y and Z are balanced. Here is a way to implement this:

```{r}
N <- 100
pop <- matrix(c(rep(0, 38 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)
```

We first rebuild our basic population, and then we use a similar strategy that we used for the social behavior. In the previous case the optimal state was to have at least one behaviour in all categories, now is to have the same amount of behaviours in the two nutrients. Thus, for the individual $1$ in the population:

```{r}
nut_y <- sum(pop[1, 17:20]) + sum(pop[1, 25:27]) + sum(pop[1, 31:32]) + pop[1, 35] + pop[1, 37]
nut_z <- sum(pop[1, 21:24]) + sum(pop[1, 28:30]) + sum(pop[1, 33:34]) + pop[1, 36] + pop[1, 38]

state <- 1 - abs(nut_y - nut_z) / 5
```

The value of $state$ is calculated by comparing how many behaviours an oranzees has that provide nutrient Y and how many the provided nutrient Z. Is then divided by 5, which is the maximum possible unbalance (all Y and no Z or viceversa, in which case $state=0$). If an oranzee, for example, has one behaviours providing Y and two providing Z, its state will be:

```{r}
state <- 1 - abs(1 - 2) / 5
state
```

Beside balancing nutrients, however, oranzees need to eat something. We can add in the $state$ calculation a part that rewards oranzees for having more food.

```{r}




The amount of food per behaviour depends from the ecological gradients we described before. Let's initialise again the environment, using as before, for the mock-up population, the values for ‘Uossob’ (remember the code for the function `set_oranzees_environment()` is in the previous document, so we call it here with the option `echo=FALSE`, meaning that is run, but not shown in the html document).

```{r echo=FALSE, message=FALSE}
library(tidyverse)

set_oranzees_environment <- function() {
  list_pop <- c("Uossob", "Iat Forest", "Ebmog", "Elaham", "Elabik", "Ognodub")
  output <- tibble(
    population = rep(list_pop, each = 38),
    behaviour = as.factor(rep(1:38, 6)),
    type = rep(c(rep("social", 16), rep("food-related", 22)), 6),
    category = rep(c(
      rep("play", 4), rep("display", 4), rep("groom", 4), rep("courthsip", 4),
      "A", "A", "A", "A", "B", "B", "B", "B", "C", "C", "C", "D", "D", "D", "E", "E", "F", "F", "G", "H", "I", "J"
    ), 6),
    nutrient = rep(c(rep(NA, 16), "Y", "Y", "Y", "Y", "Z", "Z", "Z", "Z", "Y", "Y", "Y", "Z", "Z", "Z", "Y", "Y", "Z", "Z", "Y", "Z", "Y", "Z"), 6),
    p_g = rep(NA, 38 * 6),
    p_e = rep(NA, 38 * 6)
  )

  env_or_x <- c(220, 230, 700, 705, 710, 750)
  env_or_y <- c(660, 610, 450, 430, 510, 550)

  # genetic predispositions:
  x_g <- sample(1:1000, 38)
  y_g <- sample(1:1000, 38)

  # ecological availability:
  x_e <- sample(1:1000, 38)
  y_e <- sample(1:1000, 38)

  for (i in 1:6) {
    for (behav in 1:38) {
      output[output$population == list_pop[i] & output$behaviour == behav, ]$p_g <- 1 - sqrt((x_g[behav] - env_or_x[i])^2 + (y_g[behav] - env_or_y[i])^2) / 1000

      if (behav > 16) { # only for food-related behaviours:
        output[output$population == list_pop[i] & output$behaviour == behav, ]$p_e <- 1 - sqrt((x_e[behav] - env_or_x[i])^2 + (y_e[behav] - env_or_y[i])^2) / 500
      }
    }
  }

  # replace negative values with 0s:
  if (dim(output[output$p_g < 0, ])[1] > 0) {
    output[output$p_g < 0, ]$p_g <- 0
  }
  if (dim(output[output$p_e < 0 & !is.na(output$p_e), ])[1] > 0) {
    output[output$p_e < 0 & !is.na(output$p_e), ]$p_e <- 0
  }

  # return the tibble:
  output
}
```

```{r}
oranzees_environment <- set_oranzees_environment()
test_environment <- oranzees_environment %>%
  filter(population == "Uossob")
```
  
We can think that what oranzees need to balance is not the amount of food-related behaviours per se, but the amount of food and, also, the more they eat, the better. We can rewrite the previous $state$ calculation as:

```{r}

food_behaviours <- which(pop[1, 17:38] > 0) + 16
if (length(food_behaviours) > 1) {
  amount <- test_environment$p_e[food_behaviours]
  nutrients <- test_environment$nutrient[food_behaviours]
  if (length(unique(nutrients)) > 1) {
    state <- ((1 - abs(sum(amount[nutrients == "Y"]) - sum(amount[nutrients == "Z"])) / 5) + sum(amount) / 10) / 2
  } else {
    state <- 0
  }
} else {
  state <- 0
}
```

The code is more complex: we are checking that the individual has at least two food-related beahviours (`if(length(food_behaviours) > 1)`) and that they are providing the two nutrients  (`length(unique(nutrients)) > 1`) otherwise $state=0$. If the conditions are satisfied, the state is calculated similarly to what we were doing before, but adding the amount of food per nutrients (`((1 - abs(sum(amount[nutrients == "Y"]) - sum(amount[nutrients == "Z"])) / 5)`), but there is also a second term that consider the total amount of food per se (`sum(amount / 10))`). The first is divided by 5 (the maximum possible difference between nutrients), and the second is divided by 10 (the theoretical maximum possible amount of food).

We can, as we did for social behaviours, run the calculation for all the population, together with the probability of innovating calculation `p_state <- rnorm(1, mean = 1 - state, sd = .05)`

```{r}
state <- rep(0, N)
p_state <- rep(NA, N)
for (i in 1:N) {
  food_behaviours <- which(pop[i, 17:38] > 0) + 16
  if (length(food_behaviours) > 1) {
    amount <- test_environment$p_e[food_behaviours]
    nutrients <- test_environment$nutrient[food_behaviours]
    if (length(unique(nutrients)) > 1) {
      state[i] <- ((1 - abs(sum(amount[nutrients == "Y"]) - sum(amount[nutrients == "Z"])) / 5) + sum(amount) / 10) / 2
    } else {
      state[i] <- 0
    }
  } else {
    state[i] <- 0
  }
  p_state[i] <- rnorm(1, mean = 1 - state[i], sd = .05)
}
```

With all the behaviours initalised to zero, all the values of $p_{state}$ are around one, as for social behaviours:

```{r}
head(p_state)
```

However, since the calculation of the state for food-related behaviours is more complex, is worth to do some tests to check that everything works as we want. Let's build a custom `test_environment$p_e` that we can use for the test:

```{r}
test_environment$p_e <- c(rep(NA, 16), 1, 1, 1, 1, 0, 0, 0, 0, 0, 0, 0, 1, 1, 1, 0.5, 0.5, 0.5, 0.5, 1, 0, 0, 1)
```

Category |    |    |    |    | Nutrient
---------|----|----|----|----|---------
A        | 17 **(1)**   | 18 **(1)**   | 19 **(1)** | 20 **(1)** |    Y
B        | 21 **(0)**   | 22 **(0)**   | 23 **(0)** | 24 **(0)** |    Z
C        | 25 **(0)**   | 26 **(0)**   | 27 **(0)** |            |    Y
D        | 28 **(1)**   | 29 **(1)**   | 30 **(1)** |            |    Z
E        | 31 **(0.5)** | 32 **(0.5)** |            |            |    Y
F        | 33 **(0.5)** | 34 **(0.5)** |            |           |    Z
G        | 35 **(1)**   |    |    |    |    Y
H        | 36 **(0)**   |    |    |    |    Z
I        | 37 **(0)**   |    |    |    |    Y
J        | 38 **(1)**   |    |    |    |    Z 

As well as some custom individuals in the population that cover some possible cases. We can build them looking at the table above. For example, we want an individual that has an optimal and balanced diet:

```{r}
pop[1, c(17, 28, 31, 33, 35, 38)] <- 1
```

One with an abundant, but relatively unbalanced:

```{r}
pop[2, c(17, 31, 33, 35)] <- 1
```

One with abundant, but with only one nutrient:

```{r}
pop[3, c(29, 34, 38)] <- 1
```

One with a poor but balanced diet:

```{r}
pop[4, c(31, 34)] <- 1
```

Let's rerun the calculation now (again the code is not shown), and have a look at $p_{state}$:

```{r echo=FALSE}
state <- rep(0, N)
p_state <- rep(NA, N)
for (i in 1:N) {
  food_behaviours <- which(pop[i, 17:38] > 0) + 16
  if (length(food_behaviours) > 1) {
    amount <- test_environment$p_e[food_behaviours]
    nutrients <- test_environment$nutrient[food_behaviours]
    if (length(unique(nutrients)) > 1) {
      state[i] <- ((1 - abs(sum(amount[nutrients == "Y"]) - sum(amount[nutrients == "Z"])) / 5) + sum(amount) / 10) / 2
    } else {
      state[i] <- 0
    }
  } else {
    state[i] <- 0
  }
  p_state[i] <- rnorm(1, mean = 1 - state[i], sd = .05)
}
```

```{r}
head(p_state)
```

It should show a very low value for the first individual, a somewhat middle value for the second (abundant but unbalanced), a value around one for the third (lacking a nutrient, $state=0$), again a middle value for the fourth (poor but balanced), and around one for all the rest.

***

## Population dynamics

At this point, innovations are made as they were done for social behaviours, so we can wrap everything up in the function `test_oranzees3()`. We need to change the output, to record food behaviours instead of the social ones,  

```{r}
test_oranzees3 <- function(t_max) {
  oranzees_environment <- set_oranzees_environment()
  test_environment <- oranzees_environment %>%
    filter(population == "Uossob")

  N <- 100

  pop <- matrix(c(rep(0, 38 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)

  output <- matrix(nrow = t_max, ncol = 22)

  for (t in 1:t_max) {
    output[t, ] <- colSums(pop[, 17:38])
    # demographic bit:
    pop[, 39] <- pop[, 39] + 1
    pop[pop[, 39] >= 720, ] <- 0
    old <- which(pop[, 39] > 300)
    dead <- sample(c(TRUE, FALSE), length(old), prob = c(.01, .99), replace = TRUE)
    pop[old[dead], ] <- 0
    # innovation bit:
    for (i in 1:N) {
      food_behaviours <- which(pop[i, 17:38] > 0) + 16
      if (length(food_behaviours) > 1) {
        amount <- test_environment$p_e[food_behaviours]
        nutrients <- test_environment$nutrient[food_behaviours]
        if (length(unique(nutrients)) > 1) {
          state <- ((1 - abs(sum(amount[nutrients == "Y"]) - sum(amount[nutrients == "Z"])) / 5) + sum(amount) / 10) / 2
        } else {
          state <- 0
        }
      } else {
        state <- 0
      }
      p_state <- rnorm(1, mean = 1 - state, sd = .05)
      if (runif(1) < p_state) {
        p_peering <- rnorm(22, mean = colSums(pop[, 17:38]), sd = 1)
        p_peering[p_peering < 0] <- 0
        innovation_i <- sample(17:38, 1, prob = p_peering)
        if (runif(1) < test_environment$p_g[innovation_i]) {
          if (innovation_i <= 20) {
            pop[i, 17:20] <- 0
            pop[i, innovation_i] <- 1
          } else if (innovation_i > 20 & innovation_i <= 24) {
            pop[i, 21:24] <- 0
            pop[i, innovation_i] <- 1
          } else if (innovation_i > 24 & innovation_i <= 27) {
            pop[i, 25:27] <- 0
            pop[i, innovation_i] <- 1
          } else if (innovation_i > 27 & innovation_i <= 30) {
            pop[i, 28:30] <- 0
            pop[i, innovation_i] <- 1
          } else if (innovation_i > 30 & innovation_i <= 32) {
            pop[i, 31:32] <- 0
            pop[i, innovation_i] <- 1
          } else if (innovation_i > 32 & innovation_i <= 34) {
            pop[i, 33:34] <- 0
            pop[i, innovation_i] <- 1
          } else {
            pop[i, innovation_i] <- 1
          }
        }
      }
    }
  }
  output
}
```

Here, to run:
```{r}
t_max <- 12000
my_test <- test_oranzees3(t_max)
```

And to plot the results:

```{r warning=FALSE}
my_test <- gather(as_tibble(my_test), 1:22, key = "behaviour", value = "frequency")
data_to_plot <- tibble(
  behaviour = my_test$behaviour,
  frequency = my_test$frequency,
  time = rep(1:t_max, 22),
  category = as_factor(c(
    rep("A", t_max * 4), rep("B", t_max * 4), rep("C", t_max * 3), rep("D", t_max * 3),
    rep("E", t_max * 2), rep("F", t_max * 2), rep("G", t_max), rep("H", t_max),
    rep("I", t_max), rep("J", t_max)
  ))
)

ggplot(data = data_to_plot) +
  geom_line(aes(x = time, y = frequency, color = behaviour)) +
  facet_wrap(~category) +
  theme_bw() +
  theme(legend.position = "none")
```

### Multiple runs

As we did for social behaviours, we can run the simulations several times, with the same initialisation, and look at the final results. With the same logic, here the modified function `test_oranzees4()`:

```{r}
test_oranzees4 <- function(t_max) {
  N <- 100

  pop <- matrix(c(rep(0, 38 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)

  for (t in 1:t_max) {
    # demographic bit:
    pop[, 39] <- pop[, 39] + 1
    pop[pop[, 39] >= 720, ] <- 0
    old <- which(pop[, 39] > 300)
    dead <- sample(c(TRUE, FALSE), length(old), prob = c(.01, .99), replace = TRUE)
    pop[old[dead], ] <- 0
    # innovation bit:
    for (i in 1:N) {
      food_behaviours <- which(pop[i, 17:38] > 0) + 16
      if (length(food_behaviours) > 1) {
        amount <- test_environment$p_e[food_behaviours]
        nutrients <- test_environment$nutrient[food_behaviours]
        if (length(unique(nutrients)) > 1) {
          state <- ((1 - abs(sum(amount[nutrients == "Y"]) - sum(amount[nutrients == "Z"])) / 5) + sum(amount) / 10) / 2
        } else {
          state <- 0
        }
      } else {
        state <- 0
      }
      p_state <- rnorm(1, mean = 1 - state, sd = .05)
      if (runif(1) < p_state) {
        p_peering <- rnorm(22, mean = colSums(pop[, 17:38]), sd = 1)
        p_peering[p_peering < 0] <- 0
        innovation_i <- sample(17:38, 1, prob = p_peering)
        if (runif(1) < test_environment$p_g[innovation_i]) {
          if (innovation_i <= 20) {
            pop[i, 17:20] <- 0
            pop[i, innovation_i] <- 1
          } else if (innovation_i > 20 & innovation_i <= 24) {
            pop[i, 21:24] <- 0
            pop[i, innovation_i] <- 1
          } else if (innovation_i > 24 & innovation_i <= 27) {
            pop[i, 25:27] <- 0
            pop[i, innovation_i] <- 1
          } else if (innovation_i > 27 & innovation_i <= 30) {
            pop[i, 28:30] <- 0
            pop[i, innovation_i] <- 1
          } else if (innovation_i > 30 & innovation_i <= 32) {
            pop[i, 31:32] <- 0
            pop[i, innovation_i] <- 1
          } else if (innovation_i > 32 & innovation_i <= 34) {
            pop[i, 33:34] <- 0
            pop[i, innovation_i] <- 1
          } else {
            pop[i, innovation_i] <- 1
          }
        }
      }
    }
  }
  output <- colSums(pop[, 17:38])
}
```

And here to code to run 10 replications (as for social behaviours we are not actually running it, but we saved the data once):

```{r eval=FALSE}
oranzees_environment <- set_oranzees_environment()
test_environment <- oranzees_environment %>%
  filter(population == "Uossob")

n_run <- 10
t_max <- 12000
food_test1 <- matrix(nrow = n_run, ncol = 22)
for (run in 1:n_run) {
  food_test1[run, ] <- test_oranzees4(t_max)
}
write(t(food_test1), file = "output/food_test1.csv", ncolumns = 22)
write(test_environment$p_g[17:38], file = "output/food_test1_p_g.csv", ncolumns = 1)
write(test_environment$p_e[17:38], file = "output/food_test1_p_e.csv", ncolumns = 1)
```

We can then print the results:

```{r message=FALSE}
library(reshape2)
results <- as.matrix(read.table("output/food_test1.csv"))
colnames(results) <- 17:38

as_tibble(melt(results, varnames = c("run", "behaviour"), value.name = "frequency")) %>%
  mutate(run = as_factor(run), behaviour = as_factor(behaviour)) %>%
  add_column(category = as_factor(c(
    rep("A", 40), rep("B", 40), rep("C", 30), rep("D", 30),
    rep("E", 20), rep("F", 20), rep("G", 10), rep("H", 10),
    rep("I", 10), rep("J", 10)
  ))) %>%
  ggplot() +
  geom_raster(aes(x = behaviour, y = run, fill = frequency)) +
  facet_wrap(~category, scales = "free") +
  scale_fill_gradient(low = "grey90", high = "red") +
  theme_bw()
```

### The effects of genetic propensity and ecological availability

We can then visualise the correlation between the overall frequency of a behaviour and its $p_g$ value. 

```{r}
genes <- read.table("output/food_test1_p_g.csv")
tibble(p_g = genes$V1, total = colSums(results), 
       category = as_factor(c(rep("A", 4), rep("B", 4), rep("C", 3), rep("D", 3),
    rep("E", 2), rep("F", 2), "G", "H", "I", "J"))) %>%
  ggplot(aes(y = total, x = p_g)) +
  geom_point() +
  facet_wrap(~category) +
  theme_bw()
```

And the same for its $p_e$ value:

```{r}
eco <- read.table("output/food_test1_p_e.csv")
tibble(p_e = eco$V1, total = colSums(results), 
       category = as_factor(c(rep("A", 4), rep("B", 4), rep("C", 3), rep("D", 3),
    rep("E", 2), rep("F", 2), "G", "H", "I", "J"))) %>%
  ggplot(aes(y = total, x = p_e)) +
  geom_point() +
  facet_wrap(~category) +
  theme_bw()
```

***

There are a few things to notice:

* As seen in the first figure of the 'Multiple runs' section, there is seemingly less variability with respect to social behaviours. There are many reasons why this could be the case, such as (i) more possible behaviours (10 versus 4), (ii) categories of different size (see below), or (iii) the fact that both $p_g$ and $p_e$ contribute to the results. 

* Whereas $p_g$ has a similar effect (see second figure), $p_e$ has a small effect (see third figure). In the third figure we can see, for example, that the winning behaviours in category A and B have zero $p_e$ so they do not contribute to the overall food intake. Still, they are both the ones with higher $p_g$. 

* Having categories of behaviours of different size (for social behaviours they are all categories of four behaviours) is likely to create possibly unwanted effects. In particular, as behaviours are discarded only when a new behaviour in the same category is innovated, behaviours that are the only one for their category, like the last four food-related behaviours, will be advantaged, and this is what we see here, where all four single-category behaviours are present in all runs.

***

## Additional analyses of five sets of ten runs

### Variability

To be sure that all of the above is not an idiosyncrasy of this specific set-up we run five sets of ten runs, each set with a different initial set-up, and calculate global measures (variability, effects of $p_g$ and $p_e$, and frequency of single-category behaviours). Let's first check the variability (we load the `DescTools` library, to calculate the Gini coefficient) in the first four categories (the ones with respectively four, four, three, and three behaviours):

```{r}
library(DescTools)
nset <- 5
gini_sets <- tibble(category = c(rep("A", nset), rep("B", nset), rep("C", nset), rep("D", nset)), set = as_factor(rep(1:nset,4)), gini = rep(NA, 4*nset))
for(s in 1:nset){
  results <- as.matrix(read.table(paste("output/food_test", s, ".csv", sep="")))
  gini_sets[gini_sets$set==s & gini_sets$category=="A",]$gini <- Gini(colSums(results)[1:4])
  gini_sets[gini_sets$set==s & gini_sets$category=="B",]$gini <- Gini(colSums(results)[5:8])
  gini_sets[gini_sets$set==s & gini_sets$category=="C",]$gini <- Gini(colSums(results)[9:11])
  gini_sets[gini_sets$set==s & gini_sets$category=="D",]$gini <- Gini(colSums(results)[12:14])
}
ggplot(data = gini_sets, aes(y = gini, x = category, color = category)) +
  geom_boxplot(outlier.shape = NA) + 
  geom_jitter(width = 0.2) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 1)) +
  theme(legend.position = "none")
```

Gini coefficients are quite high. We can compare the categories A and B with the results for the 10 runs of the simulations with only social behaviours:

```{r}
gini_social = rep(NA, 4)
results <- as.matrix(read.table("output/social_test1.csv"))
gini_social[1] <- Gini(colSums(results)[1:4])
gini_social[2] <-Gini(colSums(results)[5:8])
gini_social[3] <-Gini(colSums(results)[9:12])
gini_social[4] <-Gini(colSums(results)[13:16])
gini_social
```

The gini coefficients for the four categories of the simulation with only social behaviours tend to be lower (remember the comparable situation is with category A and B).

### Effects of genes and ecology

Now let's move to the effects pf $p_g$ and $p_e$. To start with, we can measure the correlation of all behaviours with their genetic propensities and ecological availability. While we loose details (the effect of $p_g$ and $p_e$ depends on the values of their 'competitors' in the same category, that is why we plotted them separately when analysing ten runs with the same initialisation), this gives an overall indication of the importance of the two factors.

```{r message=FALSE}
nset <- 5
all_runs <- tibble(frequency = rep(NA, nset * 22), p_g = rep(NA, nset * 22), p_e = rep(NA, nset * 22))
c <- 1
for(s in 1:nset){
  results <- as.matrix(read.table(paste("output/food_test", s, ".csv", sep="")))
  all_runs$frequency[c:(c+21)] <- as.numeric(colSums(results))/1000
  all_runs$p_g[c:(c+21)] <- read.table(paste("output/food_test", s, "_p_g.csv", sep=""))$V1
  all_runs$p_e[c:(c+21)] <- read.table(paste("output/food_test", s, "_p_e.csv", sep=""))$V1
  c <- c + 22
}

library(gridExtra)
p1 <- ggplot(data = all_runs) +
  geom_point(aes( x = frequency, y = p_g)) + 
  theme_bw() +
  xlim(0, 1) +
  ylim(0, 1) +
  coord_fixed()

p2 <- ggplot(data = all_runs) +
  geom_point(aes( x = frequency, y = p_e)) + 
  theme_bw() +
  xlim(0, 1) +
  ylim(0, 1) +
  coord_fixed()

grid.arrange(p1, p2, ncol = 2)
```

It looks like there is a correlation for $p_g$, but not for $p_e$. We can calculate the correlations:

```{r message=FALSE}
library(Hmisc)
corr_genes <- rcorr(all_runs$frequency, all_runs$p_g, type = "spearman")
corr_eco <- rcorr(all_runs$frequency, all_runs$p_e, type = "spearman")
```

We can look at them. For genes:
```{r}
corr_genes$r[2]
corr_genes$P[2]
```

For ecology:
```{r}
corr_eco$r[2]
corr_eco$P[2]
```

They are both correlated (which is good), still it is strange that some behaviour that bring zero food reach higher frequencies. We may want to change the $state$ calculation.

### One-behaviour categories

Finally, we can check whether there is an advantage for behaviours that are in a category without any other behaviour, i.e. categories G, H, I, J. We can compare their final frequencies with the frequencies of the behaviours in the other categories.

```{r}
tibble(type = c(rep("single_category", 4*nset), rep("others", 18*nset)), 
  frequency = c(all_runs$frequency[c(19:22, 41:44, 63:66, 85:88, 107:110)],
all_runs$frequency[-c(19:22, 41:44, 63:66, 85:88, 107:110)])) %>%
  ggplot(aes(y = frequency, x = type, color = type)) +
 # geom_boxplot(outlier.shape = NA) + 
  geom_jitter(width = 0.2) +
  theme_bw() +
  coord_cartesian(ylim = c(0, 1)) +
  theme(legend.position = "none")  
```

Results show that there are clear differences, with single-category behaviours being always successful.

***

#### TO DO

* we may want to change the $state$ calculation for food related behaviours, to give more importance to $p_e$. It is not necessary, as there is a correlation, but may be more realistic. Also it could change the results of the variability analysis, making runs more different (we do not know).

* Is the single-category behaviours a problem? If it is, there are two alternatives: (i) having ways to discard behaviours (but this will change everything, so I am not inclined), or (ii) having also for food-related behaviours categories with the same number of behaviours.

***
