---
title: "Oranzees model (03)"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R markdown file documents the development of the Oranzees model (it continues from [02-oranzees.html](02-oranzees.html)). The code-only version to run the model is in the file [main.R](main.R) in the same repository.

## Wrapping up all the code

Now that we had implemented all behaviours, we can put everything together, so to have a code to test extensively what happens in our mock-up population, before writing the simulation for the 'real' oranzees.

### Social behaviours

First, we can work to make the code more efficient. The most important thing is to vectorise it. Let's start from our first function, `test_oranzees1()`. As before, we need to run the code for the initialisation (`set_oranzees_environment()`) and for the actual function, without showing it in the html rendering (remember the option`echo=FALSE`).

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(scales)

set_oranzees_world <- function(alpha_g, alpha_e) {
  list_pop <- c("Uossob", "Iat Forest", "Ebmog", "Elaham", "Elabik", "Ognodub")
  output <- tibble(
    population = rep(list_pop, each = 64),
    behaviour = as.factor(rep(1:64, 6)),
    type = rep(c(rep("social", 32), rep("food-related", 32)), 6),
    category = rep(c(
      rep("play", 8), rep("display", 8), rep("groom", 8), rep("courthsip", 8),
      "A", "A", "A", "A", "A", "A", "A", "A", "B", "B", "B", "B", "B", "B", "B", "B","C", "C", "C", "C", "D", "D", "D", "D", "E", "E", "F", "F", "G", "H", "I", "J"
    ), 6),
    nutrient = rep(c(rep(NA, 32), "Y", "Y", "Y", "Y", "Y", "Y", "Y", "Y", "Z", "Z", "Z", "Z", "Z", "Z", "Z", "Z", "Y", "Y", "Y", "Y", "Z", "Z", "Z", "Z", "Y", "Y", "Z", "Z", "Y", "Z", "Y", "Z"), 6),
    p_g = rep(NA, 64 * 6),
    p_e = rep(NA, 64 * 6)
  )

  env_or_x <- c(220, 230, 700, 705, 710, 750)
  env_or_y <- c(660, 610, 450, 430, 510, 550)

  # genetic predispositions:
  p_g <- runif(64)
  x_g <- sample(1:1000, 64)
  y_g <- sample(1:1000, 64)
  

  # ecological availability:
  p_e <- runif(64)
  x_e <- sample(1:1000, 64)
  y_e <- sample(1:1000, 64)
  

  for(behav in 1:64){
    if(runif(1) < alpha_g){
      output[output$behaviour == behav,]$p_g <- 1 - rescale(sqrt((x_g[behav] - env_or_x)^2 + (y_g[behav] - env_or_y)^2))
    }
    else{
      output[output$behaviour == behav,]$p_g <- p_g[behav]
    }
    if(behav > 32){
      if(runif(1) < alpha_e){
        output[output$behaviour == behav,]$p_e <- 1 - rescale(sqrt((x_e[behav] - env_or_x)^2 + (y_e[behav] - env_or_y)^2))
      }
      else{
        output[output$behaviour == behav,]$p_e <- p_e[behav]
      }
    }
  }
  # return the tibble:
  output
}


test_oranzees1 <- function(t_max, alpha_g, alpha_e) {
  oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
  test_world <- oranzees_world %>%
    filter(population == "Uossob")

  N <- 100

  pop <- matrix(c(rep(0, 64 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)

  output <- matrix(nrow = t_max, ncol = 32)

  for (t in 1:t_max) {
    output[t, ] <- colSums(pop[, 1:32])
    # demographic bit:
    pop[, 65] <- pop[, 65] + 1
    pop[pop[, 65] >= 720, ] <- 0
    old <- which(pop[, 65] > 300)
    dead <- sample(c(TRUE, FALSE), length(old), prob = c(.01, .99), replace = TRUE)
    pop[old[dead], ] <- 0
    # innovation bit:
    for (i in 1:N) {
      state <- ((sum(pop[i, 1:8]) >= 1) + (sum(pop[i, 9:16]) >= 1) + (sum(pop[i, 17:24]) >= 1) + (sum(pop[i, 25:32]) >= 1)) / 4
      p_state <- rnorm(1, mean = 1 - state, sd = 0.05)
      if (runif(1) < p_state) {
        p_peering <- rnorm(32, mean = colSums(pop[, 1:32]), sd = 1)
        p_peering[p_peering < 0] <- 0
        innovation_i <- sample(1:32, 1, prob = p_peering)
        if (runif(1) < test_world$p_g[innovation_i]) {
          pop[i, innovation_i] <- 1
        }
      }
    }
  }
  output
}
```

We can now call the function. To have an idea of how efficient is the code, we can measure the time of execution. For this, we use the library `tictoc`:

```{r}
library(tictoc)
t_max <- 6000
alpha_g <- 0.1
alpha_e <- 0.9
tic()
my_test <- test_oranzees1(t_max, alpha_g, alpha_e)
toc()
```

On my laptop, the execution time is around 9 seconds. Of course, this can vary according to different hardware, what is important is not the absolute time, but the ratio between this time and the vectorised code. To vectorise the code, we need to replace most of what is iterated in the `for (i in 1:N)` cycle, applying it directly to the `pop` matrix. Let's initalise again a population, outside of the pre-written function.

```{r}
N <- 100
pop <- matrix(c(rep(0, 64 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)
```

The $state$ values can be calculated for all oranzees as:

```{r}
state <- ((rowSums(pop[, 1:8]) >= 1) + (rowSums(pop[, 9:16]) >= 1) + (rowSums(pop[, 17:24]) >= 1) + (rowSums(pop[, 25:32]) >= 1)) / 4
p_state <- rnorm(N, mean = 1 - state, sd = 0.05)
```

And we transform it in probability:

```{r}
p_state <- runif(N) < p_state
```

Now the innovation. We need to change only the calculation of `innovation_i`:

```{r}
p_peering <- rnorm(32, mean = colSums(pop[, 1:32]), sd = 1)
p_peering[p_peering < 0] <- 0
innovation_i <- sample(1:32, N, prob = p_peering, replace = TRUE)
```

And, calculate the probability to actually innovate a behaviour (given by its genetic propensity) we first need to recreate the environment: 

```{r}
oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
test_world <- oranzees_world %>%
  filter(population == "Uossob")
```

And we can then calculate it:

```{r}
p_innovate <- test_world$p_g[innovation_i]
p_innovate <- runif(N) < p_innovate * p_state
```

Putting everything together, in a new function:

```{r}
test_oranzees1_better <- function(t_max, alpha_g, alpha_e) {
  oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
  test_world<- oranzees_world %>%
    filter(population == "Uossob")

  N <- 100

  pop <- matrix(c(rep(0, 64 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)

  output <- matrix(nrow = t_max, ncol = 32)

  for (t in 1:t_max) {
    output[t, ] <- colSums(pop[, 1:32])
    # demographic bit:
    pop[, 65] <- pop[, 65] + 1
    pop[pop[, 65] >= 720, ] <- 0
    old <- which(pop[, 65] > 300)
    dead <- sample(c(TRUE, FALSE), length(old), prob = c(.01, .99), replace = TRUE)
    pop[old[dead], ] <- 0
    # innovation bit:
    state <- ((rowSums(pop[, 1:8]) >= 1) + (rowSums(pop[, 9:16]) >= 1) + (rowSums(pop[, 17:24]) >= 1) + (rowSums(pop[, 25:32]) >= 1)) / 4
    p_state <- runif(N) < rnorm(N, mean = 1 - state, sd = 0.05)
    p_peering <- rnorm(32, mean = colSums(pop[, 1:32]), sd = 1)
    p_peering[p_peering < 0] <- 0
    innovation_i <- sample(1:32, N, prob = p_peering, replace = TRUE)
    p_innovate <- runif(N) < test_world$p_g[innovation_i] * p_state
    for (i in (1:N)[p_innovate]) {
      pop[i, innovation_i[i]] <- 1
    }
    # notice the for cycle is faster than the vectorised alternative I found i.e.:
    # pop[cbind((1:N)[p_innovate], innovation_i[p_innovate])] <- 1
  }
  output
}
```

Let's test the running time, in the same way that before:

```{r}
t_max <- 6000
tic()
my_test <- test_oranzees1_better(t_max, alpha_g, alpha_e)
toc()
```

Now it is less than 2 seconds, that is, more than five time faster than before. When running many simulations these improvements are very important. 

### Food-related behaviours

Let's do the same thing for food-related behaviours. We need now to run the function `test_oranzees3()`, all the rest is already here.

```{r echo=FALSE}
test_oranzees3 <- function(t_max, alpha_g, alpha_e) {
  oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
  test_world <- oranzees_world %>%
    filter(population == "Uossob")

  N <- 100

  pop <- matrix(c(rep(0, 64 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)

  output <- matrix(nrow = t_max, ncol = 32)

  for (t in 1:t_max) {
    output[t, ] <- colSums(pop[, 33:64])
    # demographic bit:
    pop[, 65] <- pop[, 65] + 1
    pop[pop[, 65] >= 720, ] <- 0
    old <- which(pop[, 65] > 300)
    dead <- sample(c(TRUE, FALSE), length(old), prob = c(.01, .99), replace = TRUE)
    pop[old[dead], ] <- 0
    # innovation bit:
    for (i in 1:N) {
      nut_y <- (sum(pop[1, 33:40])>=1) + (sum(pop[1, 49:52])>=1) + (sum(pop[1, 57:58])>=1) + pop[1, 61] + pop[1, 63]
      nut_z <- (sum(pop[1, 41:48])>=1) + (sum(pop[1, 53:56])>=1) + (sum(pop[1, 59:60])>=1) + pop[1, 62] + pop[1, 64]
      state <- (nut_y + nut_z - abs(nut_y - nut_z)) / 10
      p_state <- rnorm(1, mean = 1 - state, sd = .05)
      if (runif(1) < p_state) {
        p_peering <- rnorm(32, mean = colSums(pop[, 33:64]), sd = 1)
        p_peering[p_peering < 0] <- 0
        innovation_i <- sample(33:64, 1, prob = p_peering)
        if (runif(1) < (test_world$p_g[innovation_i] * test_world$p_e[innovation_i])) {
          pop[i, innovation_i] <- 1
        }
      }
    }
  }
  output
}
```

Here is the running time:

```{r}
t_max <- 6000
tic()
my_test <- test_oranzees3(t_max, alpha_g, alpha_e)
toc()
```

The logic is the same of what we just did for social behaviours, so here is directly the new function:

```{r}
test_oranzees3_better <- function(t_max, alpha_g, alpha_e) {
  oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
  test_world<- oranzees_world %>%
    filter(population == "Uossob")

  N <- 100

  pop <- matrix(c(rep(0, 65 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)

  output <- matrix(nrow = t_max, ncol = 32)

  for (t in 1:t_max) {
    output[t, ] <- colSums(pop[, 33:64])
    # demographic bit:
    pop[, 65] <- pop[, 65] + 1
    pop[pop[, 65] >= 720, ] <- 0
    old <- which(pop[, 65] > 300)
    dead <- sample(c(TRUE, FALSE), length(old), prob = c(.01, .99), replace = TRUE)
    pop[old[dead], ] <- 0
    # innovation bit:
    nut_y <- (rowSums(pop[, 33:40])>=1) + (rowSums(pop[, 49:52])>=1) + (rowSums(pop[, 57:58])>=1) + pop[, 61] + pop[, 64]
    nut_z <- (rowSums(pop[, 41:48])>=1) + (rowSums(pop[, 53:56])>=1) + (rowSums(pop[, 59:60])>=1) + pop[, 62] + pop[, 63]
    state <- (nut_y + nut_z - abs(nut_y - nut_z)) / 10
    p_state <- runif(N) < rnorm(N, mean = 1 - state, sd = 0.05)
    p_peering <- rnorm(32, mean = colSums(pop[, 33:64]), sd = 1)
    p_peering[p_peering < 0] <- 0
    innovation_i <- sample(33:64, N, prob = p_peering, replace = TRUE)
    p_innovate <- runif(N) < test_world$p_g[innovation_i] * test_world$p_e[innovation_i] * p_state
    for (i in (1:N)[p_innovate]) {
      pop[i, innovation_i[i]] <- 1
    }
  }
  output
}
```

We can test the running time:

```{r}
t_max <- 6000
tic()
my_test <- test_oranzees3_better(t_max, alpha_g, alpha_e)
toc()
```

This is usually between 20 and 30 times faster! 

***

## Final code for mock-up population

It is finally time to put all together, for a simulation running single runs of the complete (social plus food-related behaviours). We will use this to perform some in-depth analyses to be sure that everything works as intended, before applying it to the 'real' oranzees populations. The final code make use of the function we already wrote, `set_oranzees_world()`. We need to rewrite the optimised versions of the social and the food-related behaviours, taking out global instructions (e.g. demography, or the initalisation of the environment) plus a final function that puts everything together. It is easy to start from how the final function should look like, and then add the sub-functions.

```{r}
mockup_oranzees <- function(t_max, alpha_g, alpha_e, init_world) {
  # initalise everything:
  N <- 100
  pop <- matrix(c(rep(0, 64 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)
  output <- matrix(nrow = t_max, ncol = 64)
  if (init_world) {
    oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
    test_world <- oranzees_world %>%
      filter(population == "Uossob")
  }
  # start simulation here:
  for (t in 1:t_max) {
    output[t,] <- colSums(pop[, 1:64])
    pop <- update_demography(pop)
    pop <- update_social_behaviours(pop, test_world)
    pop <- update_food_behaviours(pop, test_world)
  }
  output
}
```

Something like that should do what we want. Notice that we still need to write the appropriate functions, and notice also the new parameter `init_world`. We will pass `TRUE` (or $1$) if we want the environment to be initialised each time we run the function, and `FALSE` (or $0$) if we want to use an environment created separately. This is helpful to run multiple tests with the same or different environments (before we were using two different functions).

Let's write here the three new functions we need:

```{r}
update_demography <- function(pop) {
  pop[, 65] <- pop[, 65] + 1
  pop[pop[, 65] >= 720, ] <- 0
  old <- which(pop[, 65] > 300)
  dead <- sample(c(TRUE, FALSE), length(old), prob = c(.01, .99), replace = TRUE)
  pop[old[dead], ] <- 0
  pop
}

update_social_behaviours <- function(pop, test_world) {
  N <- dim(pop)[1]
  state <- ((rowSums(pop[, 1:8]) >= 1) + (rowSums(pop[, 9:16]) >= 1) + (rowSums(pop[, 17:24]) >= 1) + (rowSums(pop[, 25:32]) >= 1)) / 4
  p_state <- runif(N) < rnorm(N, mean = 1 - state, sd = 0.05)
  p_peering <- rnorm(32, mean = colSums(pop[, 1:32]), sd = 1)
  p_peering[p_peering < 0] <- 0
  innovation_i <- sample(1:32, N, prob = p_peering, replace = TRUE)
  p_innovate <- runif(N) < test_world$p_g[innovation_i]  * p_state
  for (i in (1:N)[p_innovate]) {
    pop[i, innovation_i[i]] <- 1
  }
  pop
}  

update_food_behaviours <- function(pop, test_world) {
  N <- dim(pop)[1]
  nut_y <- (rowSums(pop[, 33:40])>=1) + (rowSums(pop[, 49:52])>=1) + (rowSums(pop[, 57:58])>=1) + pop[, 61] + pop[, 64]
  nut_z <- (rowSums(pop[, 41:48])>=1) + (rowSums(pop[, 53:56])>=1) + (rowSums(pop[, 59:60])>=1) + pop[, 62] + pop[, 63]
  state <- (nut_y + nut_z - abs(nut_y - nut_z)) / 10 
  p_state <- runif(N) < rnorm(N, mean = 1 - state, sd = 0.05)
  p_peering <- rnorm(32, mean = colSums(pop[, 33:64]), sd = 1)
  p_peering[p_peering < 0] <- 0
  innovation_i <- sample(33:64, N, prob = p_peering, replace = TRUE)
  p_innovate <- runif(N) < test_world$p_g[innovation_i] * test_world$p_e[innovation_i] * p_state
  for (i in (1:N)[p_innovate]) {
    pop[i, innovation_i[i]] <- 1
  }
  pop
}
```

We can try and run it now:

```{r}
t_max <- 6000
my_test <- mockup_oranzees(t_max, alpha_g, alpha_e, TRUE)
```

And visualise the results:

```{r warning = FALSE}
my_test <- gather(as_tibble(my_test), 1:64, key = "behaviour", value = "frequency")
data_to_plot <- tibble(
  behaviour = my_test$behaviour,
  frequency = my_test$frequency,
  time = rep(1:t_max, 64),
  category = as_factor(c(
    rep("play", t_max * 8), rep("display", t_max * 8), rep("groom", t_max * 8), rep("courthsip", t_max * 8),
    rep("A", t_max * 8), rep("B", t_max * 8), rep("C", t_max * 4), rep("D", t_max * 4),
    rep("E", t_max * 2), rep("F", t_max * 2), rep("G", t_max), rep("H", t_max),
    rep("I", t_max), rep("J", t_max)
  ))
)

ggplot(data = data_to_plot) +
  geom_line(aes(x = time, y = frequency, color = behaviour)) +
  facet_wrap(~category) +
  theme_bw() +
  theme(legend.position = "none")
```

Just to check that everything work as expected, we can run, as we did before separately for social and food-related behaviours, 10 runs with the same initalisation and visualise the results. We can rewrite a the function with the additional parameter $n_run$. If $n_run>1$ the fucntion does not store all the history of the traits, but only the final state of the population.

```{r}
mockup_oranzees <- function(t_max, alpha_g, alpha_e, init_world, n_run) {
  
  N <- 100
  
  if(n_run == 1){
    output <- matrix(nrow = t_max, ncol = 64)
  }
  else{
    output <- matrix(nrow = n_run, ncol = 64)
  }
  
  oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
  test_world <- oranzees_world %>%
    filter(population == "Uossob")
  
  for(run in 1:n_run){
    pop <- matrix(c(rep(0, 64 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)
    if (init_world) {
      oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
      test_world <- oranzees_world %>%
        filter(population == "Uossob")
    }
    # start simulation here:
    for (t in 1:t_max) {
      if(n_run == 1){
        output[t,] <- colSums(pop[, 1:64])
      }
      pop <- update_demography(pop)
      pop <- update_social_behaviours(pop, test_world)
      pop <- update_food_behaviours(pop, test_world)
    }
    if( n_run > 1){
      output[run, ] <- colSums(pop[, 1:64])
    }  
  }
  output
}
```

It is now easy to run the simulation online. To repeat the same test, we simply need to call:

```{r}
my_test <- mockup_oranzees(t_max = 6000, alpha_g = 0.1, alpha_e = 0.9, init_world = FALSE, n_run = 10)
```

And plot:

```{r message=FALSE, fig.height = 7}
library(reshape2)
as_tibble(melt(my_test, varnames = c("run", "behaviour"), value.name = "frequency")) %>%
  mutate(run = as_factor(run), behaviour = as_factor(behaviour)) %>%
  add_column(category = as_factor(c(
    rep("play", 80), rep("display", 80), rep("groom", 80), rep("courthsip", 80),
    rep("A", 80), rep("B", 80), rep("C", 40), rep("D", 40),
    rep("E", 20), rep("F", 20), rep("G", 10), rep("H", 10),
    rep("I", 10), rep("J", 10)
  ))) %>%
  ggplot() +
  geom_raster(aes(x = behaviour, y = run, fill = frequency)) +
  facet_wrap(~category, scales = "free") +
  scale_fill_gradient(low = "grey90", high = "red") +
  theme_bw()
```

***

The documentation continues in [04-oranzees.html](04-oranzees.html)

