---
title: "Oranzees model (04)"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R markdown file documents the development of the Oranzees model (it continues from [03-oranzees.html](03-oranzees.html)). The code-only version to run the model is in the file [main.R](main.R) in the same repository.

### Tests at individual-level

So far, we focused on the results at population and trait-level: how many traits are in the population? What are the successful traits? And so on. Another important aspect is to understand what happens at individual level. Are there some individuals with many traits and other with few? What is the effect of age? How do these features interact with our other parameters, such as the two values of $\alpha$?

We first run the code for the last versions of the functions we wrote, without showing it in the html rendering (remember the option`echo=FALSE`).

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(scales)

set_oranzees_world <- function(alpha_g, alpha_e) {
  list_pop <- c("Uossob", "Iat Forest", "Ebmog", "Elaham", "Elabik", "Ognodub")
  output <- tibble(
    population = rep(list_pop, each = 64),
    behaviour = as.factor(rep(1:64, 6)),
    type = rep(c(rep("social", 32), rep("food-related", 32)), 6),
    category = rep(c(
      rep("play", 8), rep("display", 8), rep("groom", 8), rep("courthsip", 8),
      "A", "A", "A", "A", "A", "A", "A", "A", "B", "B", "B", "B", "B", "B", "B", "B","C", "C", "C", "C", "D", "D", "D", "D", "E", "E", "F", "F", "G", "H", "I", "J"
    ), 6),
    nutrient = rep(c(rep(NA, 32), "Y", "Y", "Y", "Y", "Y", "Y", "Y", "Y", "Z", "Z", "Z", "Z", "Z", "Z", "Z", "Z", "Y", "Y", "Y", "Y", "Z", "Z", "Z", "Z", "Y", "Y", "Z", "Z", "Y", "Z", "Y", "Z"), 6),
    p_g = rep(NA, 64 * 6),
    p_e = rep(NA, 64 * 6)
  )

  env_or_x <- c(220, 230, 700, 705, 710, 750)
  env_or_y <- c(660, 610, 450, 430, 510, 550)

  # genetic predispositions:
  p_g <- runif(64)
  x_g <- sample(1:1000, 64)
  y_g <- sample(1:1000, 64)
  

  # ecological availability:
  p_e <- runif(64)
  x_e <- sample(1:1000, 64)
  y_e <- sample(1:1000, 64)
  

  for(behav in 1:64){
    if(runif(1) < alpha_g){
      output[output$behaviour == behav,]$p_g <- 1 - rescale(sqrt((x_g[behav] - env_or_x)^2 + (y_g[behav] - env_or_y)^2))
    }
    else{
      output[output$behaviour == behav,]$p_g <- p_g[behav]
    }
    if(behav > 32){
      if(runif(1) < alpha_e){
        output[output$behaviour == behav,]$p_e <- 1 - rescale(sqrt((x_e[behav] - env_or_x)^2 + (y_e[behav] - env_or_y)^2))
      }
      else{
        output[output$behaviour == behav,]$p_e <- p_e[behav]
      }
    }
  }
  # return the tibble:
  output
}

update_demography <- function(pop) {
  pop[, 65] <- pop[, 65] + 1
  pop[pop[, 65] >= 720, ] <- 0
  old <- which(pop[, 65] > 300)
  dead <- sample(c(TRUE, FALSE), length(old), prob = c(.01, .99), replace = TRUE)
  pop[old[dead], ] <- 0
  pop
}

update_social_behaviours <- function(pop, test_world) {
  N <- dim(pop)[1]
  state <- ((rowSums(pop[, 1:8]) >= 1) + (rowSums(pop[, 9:16]) >= 1) + (rowSums(pop[, 17:24]) >= 1) + (rowSums(pop[, 25:32]) >= 1)) / 4
  p_state <- runif(N) < rnorm(N, mean = 1 - state, sd = 0.05)
  p_peering <- rnorm(32, mean = colSums(pop[, 1:32]), sd = 1)
  p_peering[p_peering < 0] <- 0
  innovation_i <- sample(1:32, N, prob = p_peering, replace = TRUE)
  p_innovate <- runif(N) < test_world$p_g[innovation_i]  * p_state
  for (i in (1:N)[p_innovate]) {
    pop[i, innovation_i[i]] <- 1
  }
  pop
}  

update_food_behaviours <- function(pop, test_world) {
  N <- dim(pop)[1]
  nut_y <- (rowSums(pop[, 33:40])>=1) + (rowSums(pop[, 49:52])>=1) + (rowSums(pop[, 57:58])>=1) + pop[, 61] + pop[, 64]
  nut_z <- (rowSums(pop[, 41:48])>=1) + (rowSums(pop[, 53:56])>=1) + (rowSums(pop[, 59:60])>=1) + pop[, 62] + pop[, 63]
  state <- (nut_y + nut_z - abs(nut_y - nut_z)) / 10 
  p_state <- runif(N) < rnorm(N, mean = 1 - state, sd = 0.05)
  p_peering <- rnorm(32, mean = colSums(pop[, 33:64]), sd = 1)
  p_peering[p_peering < 0] <- 0
  innovation_i <- sample(33:64, N, prob = p_peering, replace = TRUE)
  p_innovate <- runif(N) < test_world$p_g[innovation_i] * test_world$p_e[innovation_i] * p_state
  for (i in (1:N)[p_innovate]) {
    pop[i, innovation_i[i]] <- 1
  }
  pop
}
```

However, we need to do some modifications to the main function `mockup_oranzees()`, as we want now, as output, the final composition of the population, and not the frequency of the traits:

```{r}
mockup_oranzees_pop <- function(t_max, alpha_g, alpha_e) {
  N <- 100
  oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
  test_world <- oranzees_world %>%
    filter(population == "Uossob")
  pop <- matrix(c(rep(0, 64 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)
  # start simulation here:
  for (t in 1:t_max) {
    pop <- update_demography(pop)
    pop <- update_social_behaviours(pop, test_world)
    pop <- update_food_behaviours(pop, test_world)
  }
  output <- pop
}
```

The new function has less parameters than `mockup_oranzees()`, since it only runs by default a single test, and it gives as output all the population. We can run it, starting with the 'usual' parameter values:

```{r}
my_test <- mockup_oranzees_pop(t_max = 6000, alpha_g = 0.1, alpha_e = 0.9)
```

We can write `head(my_test)` (remember it is a matrix), to see the first lines of the output:

```{r}
head(my_test)
```

Each of the first six lines shows an oranzee. A first aspect is, as said before, the distribution of traits among the population. 

```{r}
tibble(sum_traits = rowSums(my_test[, 1:64]), individual = 1:100) %>%
  ggplot() +
  geom_histogram(aes(x = sum_traits), bins = 15, colour = "black") +
  labs(x = "Number of traits", y = "Number of oranzees") +
  geom_vline(aes(xintercept = mean(sum_traits)), color = "red") +
  theme_bw()
```

The distribution looks as expected. Oranzees have an average number of traits of 15 or slightly more (this changes depending on the run), and the distribution is roughly normal. The left side of the distribution is however more skewed, meaning that some individuals have less traits that what would be predicted. They are likely to be younger individuals, that did not have time yet to acquire their traits. To check this, we can visually check the correlation between age and number of traits:

```{r}
tibble(sum_traits = rowSums(my_test[, 1:64]), individual = 1:100, age = my_test[, 65]) %>%
  ggplot(aes(x = age / 12, y = sum_traits)) +
  geom_point() +
  labs(x = "Age (in years)", y = "Number of traits") +
  geom_smooth(method = loess) +
  theme_bw()
```

This plot shows that, indeed, younger individuals have less traits, and that they acquire them during their life. Not only, it also shows that, at a certain age, usually between 10 and 20 years, the number of traits arrive at a plateau, meaning that oranzees stop to acquire new traits. 

Another way to look at similar information is to follow each individual through time. We can modify again `mockup_oranzee()` to produce as output the full story of all oranzees stored in one slot of the matrix population, say the first one (this choice is purely arbitrary). Since when an oranzee dies a naive one replace him in the matrix, a slot in a matrix is effectively a random sample of the population:

```{r}
mockup_oranzees_pop2 <- function(t_max, alpha_g, alpha_e) {
  N <- 100
  output <- matrix(nrow = t_max, ncol = 65)
  oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
  test_world <- oranzees_world %>%
    filter(population == "Uossob")
  pop <- matrix(c(rep(0, 64 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)
  # start simulation here:
  for (t in 1:t_max) {
    pop <- update_demography(pop)
    pop <- update_social_behaviours(pop, test_world)
    pop <- update_food_behaviours(pop, test_world)
    output[t, ] <- pop[1, ]
  }
  output
}
```

Again, let's run if for $\alpha_g=0.1$ and $\alpha_e=0.9$:

```{r}
my_test <- mockup_oranzees_pop2(t_max = 6000, alpha_e = 0.9, alpha_g = 0.1)
```

We can check how many individuals are in our sample.

```{r}
sum(my_test[, 65] == 0)
```

And at which age they died:

```{r}
my_test[which(my_test[, 65] == 0) - 1, 65] / 12
```

More interestingly, we can check how they acquired new traits through their life time:

```{r}
t_max <- 6000
newborn_time <- which(my_test[, 65] == 0)
individual <- as_factor(rep(
  seq(1:(sum(my_test[, 65] == 0) + 1)),
  c(newborn_time[1] - 1, diff(newborn_time), t_max + 1 - last(newborn_time))
))

tibble(n_traits = rowSums(my_test[, 1:64]), age = my_test[, 65] / 12, oranzee_ID = individual) %>%
  ggplot(aes(x = age, y = n_traits, colour = oranzee_ID)) +
  geom_line() +
  theme_bw()
```

As expected. Notice there should be an oranzee (the one with $ID=1$) that starts somewhere in the middle of the x-axis. This is the oranzee in the initial population, where all individuals are initialised with random ages.

***

The documentation continues in [05-oranzees.html](05-oranzees.html)


