---
title: "Oranzees model (05)"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R markdown file documents the development of the Oranzees model (it continues from [04-oranzees.html](04-oranzees.html)). The code-only version to run the model is in the file [main.R](main.R) in the same repository.

### Whiten et al. codes for behaviours

In the Whiten et al., 1999 paper there are not clear quantitative definitions, but we can test the codes as follow:

* *customary* is a behaviour observed in over 50% of individuals in at least one age class.

* *habitual* is a behaviour observed in at least two individuals over all the population (CT suggest 'observed more than once', but we do not have actual performance of the behaviours in the model).

* *present* is a behaviour observed in at least one individual over all the population.

* *absent* is a behviour never observed.

* *ecological explanations* is a behaviour that is absent because of local ecological features.

Notice the last category in Whiten et al., 1999 (*unknown*, i.e. "the behaviour has not been recorded, but this may be due to inadequacy of relevant observational opportunities") does not apply in our case.

To use these codes with our simulations we need to modify again the function `mockup_oranzees()`, as we need a richer output. We not only need the number of behaviours, but we need to know which individual was performing them (so to have its age), as well as the values of $p_e$ (ecological availability) for the food behaviours, so to determine if an *ecological explanation* applies. In fact, the best way of proceeding is to write a further function so that it gives directly as output the codes we need. Let's call the new function `mockup_oranzees_codes()`:

```{r}
mockup_oranzees_codes <- function(t_max, alpha_g, alpha_e, init_world, n_run) {
  
  N <- 100

  output <- matrix(nrow = n_run, ncol = 5)

  oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
  test_world <- oranzees_world %>%
    filter(population == "Uossob")

  for(run in 1:n_run){
    pop <- matrix(c(rep(0, 64 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)
    if (init_world) {
      oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
      test_world <- oranzees_world %>%
        filter(population == "Uossob")
    }
    # start simulation here:
    for (t in 1:t_max) {
      pop <- update_demography(pop)
      pop <- update_social_behaviours(pop, test_world)
      pop <- update_food_behaviours(pop, test_world)
    }
    # calculate codes values:

    # age classes:
    adults = pop[,65]/12 > 16
    subadults = pop[,65]/12 > 8 & pop[,65]/12 <= 16
    juveniles = pop[,65]/12 <= 8

    # customary:
    customary_adults <- rep(FALSE, 64)
    if(sum(adults) >= 3){
      customary_adults <- colSums(pop[adults,1:64])>(sum(adults)/2)
    }
    customary_subadults <- rep(FALSE, 64)
    if(sum(subadults) >= 3){
      customary_subadults <- colSums(pop[subadults,1:64])>(sum(subadults)/2)
    }
    customary_juveniles <- rep(FALSE, 64)
    if(sum(juveniles) >= 3){
      customary_juveniles <- colSums(pop[juveniles,1:64])>(sum(juveniles)/2)
    }
    customary <- customary_adults | customary_subadults | customary_juveniles
    output[run, 1] <- sum(customary)

    # habitual:
    habitual <- colSums(pop[,1:64])>=2 & !customary
    output[run, 2] <- sum(habitual)

    # present:
    present <- colSums(pop[,1:64])==1
    output[run, 3] <- sum(present)

    # absent or ecological explanation:
    all_absent <- !(customary | habitual | present)
    absent <- all_absent & (test_world$p_e > 0 | test_world$type == "social")
    output[run, 4] <- sum(absent)
    ecological_explanation <- all_absent & (test_world$p_e == 0 & test_world$type == "food-related" )
    output[run, 5] <- sum(ecological_explanation)
  }
  output
}
```

The novel part follows the comment `# calculate codes values`. First, we divide the population in three age classes (*adults* are more than 16 year old, *subadults* between 8 and 16, and *juveniles* less than 8 year old). We can then calcualte the actual codes.

Let's run 10 simulations in a "standard" situation, with a stronger effect of ecological availability with respect to genetic propensity ($\alpha_g=0.1$ and $\alpha_e=0.9$). We first need to call all the functions we already wrote, except the original `mockup_oranzees()` (code not shown). 

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(scales)
library(reshape2)
library(viridis)

#############################
# COMMON FUNCTIONS:
#############################

set_oranzees_world <- function(alpha_g, alpha_e) {
  list_pop <- c("Uossob", "Iat Forest", "Ebmog", "Elaham", "Elabik", "Ognodub")
  output <- tibble(
    population = rep(list_pop, each = 64),
    behaviour = as.factor(rep(1:64, 6)),
    type = rep(c(rep("social", 32), rep("food-related", 32)), 6),
    category = rep(c(
      rep("play", 8), rep("display", 8), rep("groom", 8), rep("courthsip", 8),
      "A", "A", "A", "A", "A", "A", "A", "A", "B", "B", "B", "B", "B", "B", "B", "B","C", "C", "C", "C", "D", "D", "D", "D", "E", "E", "F", "F", "G", "H", "I", "J"
    ), 6),
    nutrient = rep(c(rep(NA, 32), "Y", "Y", "Y", "Y", "Y", "Y", "Y", "Y", "Z", "Z", "Z", "Z", "Z", "Z", "Z", "Z", "Y", "Y", "Y", "Y", "Z", "Z", "Z", "Z", "Y", "Y", "Z", "Z", "Y", "Z", "Y", "Z"), 6),
    p_g = rep(NA, 64 * 6),
    p_e = rep(NA, 64 * 6)
  )

  env_or_x <- c(220, 230, 700, 705, 710, 750)
  env_or_y <- c(660, 610, 450, 430, 510, 550)

  # genetic predispositions:
  p_g <- runif(64)
  x_g <- sample(1:1000, 64)
  y_g <- sample(1:1000, 64)
  

  # ecological availability:
  p_e <- runif(64)
  x_e <- sample(1:1000, 64)
  y_e <- sample(1:1000, 64)
  

  for(behav in 1:64){
    if(runif(1) < alpha_g){
      output[output$behaviour == behav,]$p_g <- 1 - rescale(sqrt((x_g[behav] - env_or_x)^2 + (y_g[behav] - env_or_y)^2))
    }
    else{
      output[output$behaviour == behav,]$p_g <- p_g[behav]
    }
    if(behav > 32){
      if(runif(1) < alpha_e){
        output[output$behaviour == behav,]$p_e <- 1 - rescale(sqrt((x_e[behav] - env_or_x)^2 + (y_e[behav] - env_or_y)^2))
      }
      else{
        output[output$behaviour == behav,]$p_e <- p_e[behav]
      }
    }
  }
  # return the tibble:
  output
}

update_demography <- function(pop) {
  pop[, 65] <- pop[, 65] + 1
  pop[pop[, 65] >= 720, ] <- 0
  old <- which(pop[, 65] > 300)
  dead <- sample(c(TRUE, FALSE), length(old), prob = c(.01, .99), replace = TRUE)
  pop[old[dead], ] <- 0
  pop
}

update_social_behaviours <- function(pop, test_world) {
  N <- dim(pop)[1]
  state <- ((rowSums(pop[, 1:8]) >= 1) + (rowSums(pop[, 9:16]) >= 1) + (rowSums(pop[, 17:24]) >= 1) + (rowSums(pop[, 25:32]) >= 1)) / 4
  p_state <- runif(N) < rnorm(N, mean = 1 - state, sd = 0.05)
  p_peering <- rnorm(32, mean = colSums(pop[, 1:32]), sd = 1)
  p_peering[p_peering < 0] <- 0
  innovation_i <- sample(1:32, N, prob = p_peering, replace = TRUE)
  p_innovate <- runif(N) < test_world$p_g[innovation_i]  * p_state
  for (i in (1:N)[p_innovate]) {
    pop[i, innovation_i[i]] <- 1
  }
  pop
}  

update_food_behaviours <- function(pop, test_world) {
  N <- dim(pop)[1]
  nut_y <- (rowSums(pop[, 33:40])>=1) + (rowSums(pop[, 49:52])>=1) + (rowSums(pop[, 57:58])>=1) + pop[, 61] + pop[, 64]
  nut_z <- (rowSums(pop[, 41:48])>=1) + (rowSums(pop[, 53:56])>=1) + (rowSums(pop[, 59:60])>=1) + pop[, 62] + pop[, 63]
  state <- (nut_y + nut_z - abs(nut_y - nut_z)) / 10 
  p_state <- runif(N) < rnorm(N, mean = 1 - state, sd = 0.05)
  p_peering <- rnorm(32, mean = colSums(pop[, 33:64]), sd = 1)
  p_peering[p_peering < 0] <- 0
  innovation_i <- sample(33:64, N, prob = p_peering, replace = TRUE)
  p_innovate <- runif(N) < test_world$p_g[innovation_i] * test_world$p_e[innovation_i] * p_state
  for (i in (1:N)[p_innovate]) {
    pop[i, innovation_i[i]] <- 1
  }
  pop
}
```

We then call the new `mockup_oranzees_codes()`:

```{r}
my_test <- mockup_oranzees_codes(t_max = 6000, alpha_g = 0.1, alpha_e = 0.9, init_world = TRUE, n_run = 10)
```

Let's plot the results:

```{r}
tibble(code = as_factor(rep(c("customary", "habitual", "present", "absent", "ecological"), each=10)),
       behaviours = as.vector(my_test)) %>%
  ggplot(aes( x = code, y = behaviours, fill = code)) +
    geom_boxplot() +
    geom_jitter(width=0.05, alpha=0.5) +
    theme_bw()
```


### The six populations

Up until now, we worked on a sinlge population to test the model. The ultimate goal of the model is, however, to compare the "cultures" in the six populations of oranzees (remember the function `set_oranzees_world()`). 

We should now write a new function `run_oranzees()` that iterates the simulation on the six populations and outputs the results we are interested in. For now, we use the various implementation choices made above, and we set six custom population sizes.

```{r}
run_oranzees <- function(t_max, alpha_g, alpha_e, init_world, n_run) {
  
  N <- c(50, 100, 150, 50, 100, 150)
  
  oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
  
  output <- tibble(run = rep(1:n_run, each= 64 * 6),
                    population = rep(oranzees_world$population, n_run),
                   behaviour = rep(1:64, 6 * n_run),
                    code = rep("NA", 64 * 6 * n_run))
  
  for(run in 1:n_run){
    
    if (init_world) {
      oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
    }
    
    for(current_population in 1:6){ 
      pop <- matrix(c(rep(0, 64 * N[current_population]), 
                      sample(1:300, N[current_population], replace = TRUE)), 
                    nrow = N[current_population], byrow = FALSE)
      current_world <- oranzees_world %>%
        filter(population == unique(oranzees_world$population)[current_population])
      # start simulation here:
      for (t in 1:t_max) {
        pop <- update_demography(pop)
        pop <- update_social_behaviours(pop, current_world)
        pop <- update_food_behaviours(pop, current_world)
      }
      
      # calculate output:
    
      # age classes:
      adults = pop[,65]/12 > 16
      subadults = pop[,65]/12 > 8 & pop[,65]/12 <= 16
      juveniles = pop[,65]/12 <= 8
    
      # customary:

      customary_adults <- rep(FALSE, 64)
      if(sum(adults) >= 3){
        customary_adults <- colSums(pop[adults,1:64])>(sum(adults)/2) 
      }
      customary_subadults <- rep(FALSE, 64)
      if(sum(subadults) >= 3){
        customary_subadults <- colSums(pop[subadults,1:64])>(sum(subadults)/2)
      }
      customary_juveniles <- rep(FALSE, 64)
      if(sum(juveniles) >= 3){
        customary_juveniles <- colSums(pop[juveniles,1:64])>(sum(juveniles)/2) 
      }
      customary <- customary_adults | customary_subadults | customary_juveniles
      output[output$population==current_world$population & output$run == run & customary==TRUE,4] <- "customary"
    
      # habitual:
      habitual <- colSums(pop[,1:64])>=2 & !customary
      output[output$population==current_world$population & output$run == run & habitual==TRUE,4] <- "habitual"
      
      # present:
      present <- colSums(pop[,1:64])==1
      output[output$population==current_world$population & output$run == run & present==TRUE,4] <- "present"
      
      # absent or ecological explanation:
      all_absent <- !(customary | habitual | present)
      absent <- all_absent & (current_world$type == "social" | current_world$p_e > 0)
      output[output$population==current_world$population & output$run == run & absent==TRUE,4] <- "absent"
      ecological_explanation <- all_absent & (current_world$type == "food-related" & current_world$p_e == 0)
      output[output$population==current_world$population & output$run == run & ecological_explanation==TRUE,4] <- "ecological explanation"
    }
  }
  output
}

```

The function is similar to the previous `mockup_oranzees_codes()` with few key differences:

* we set *N* for the six populations, with: `N <- c(50, 100, 150, 50, 100, 150)` 

* the `output` is different (we will explore it later)

* the simulation is run six times, one for each population.

We can now run it:

```{r}
my_test <- run_oranzees(t_max = 6000, alpha_g = 0.1, alpha_e = 0.9, init_world = TRUE, n_run = 1)
```

Let's have a look at the output:

```{r}
my_test
```

Each line represents a behaviour, in each population, in each run. Now we have only one run, and we can visualise the results. We divide the behaviour in "social" and "food-related" in two different plots so they are easier to read. Notice the first two lines of the code: they are needed to position the populations in the "right" order (otherwise `ggplot` would position them in alphabetical order). Let's plot the social behaviours first:

```{r message = FALSE, fig.height = 14}
library(viridis)
my_test$population <- as.factor(my_test$population)
levels(my_test$population) <- unique(my_test$population) 
# social behaviour
my_test %>%
  filter(behaviour <= 32) %>%
  ggplot( aes(x = population, y = behaviour, colour = code)) +
  geom_point(size = 10) +
  scale_color_viridis(discrete = TRUE, option = "D") +
  ggtitle("Social behaviours") +
  theme_bw()
```

And the food-related behaviours:

```{r, fig.height = 14}
my_test %>%
  filter(behaviour > 32) %>%
  ggplot( aes(x = population, y = behaviour, colour = code)) +
  geom_point(size = 10) +
  scale_color_viridis(discrete = TRUE, option = "D") +
  ggtitle("Food-related behaviours") +
  theme_bw()
```

***

The documentation continues in [06-oranzees.html](06-oranzees.html) 



