---
title: "Oranzees model (05)"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R markdown file documents the development of the Oranzees model (it continues from [04-oranzees.html](04-oranzees.html)). The code-only version to run the model is in the file [main.R](main.R) in the same repository.

### Tests at individual-level

So far, we focused on the results at population and trait-level: how many traits are in the population? What are the successfull traits? And so on. Another important aspect is to understand what happens at individual level. Are there some individuals with many traits and other with few? What is the effect of age? How do these features interact with our other parameters, such as the two values of $\alpha$?

As usual, We first run the code for the last versions of the functions we wrote, without showing it in the html rendering (remember the option`echo=FALSE`).

```{r echo=FALSE, message=FALSE}
library(tidyverse)
library(scales)

set_oranzees_world <- function(alpha_g, alpha_e) {
  list_pop <- c("Uossob", "Iat Forest", "Ebmog", "Elaham", "Elabik", "Ognodub")
  output <- tibble(
    population = rep(list_pop, each = 38),
    behaviour = as.factor(rep(1:38, 6)),
    type = rep(c(rep("social", 16), rep("food-related", 22)), 6),
    category = rep(c(
      rep("play", 4), rep("display", 4), rep("groom", 4), rep("courthsip", 4),
      "A", "A", "A", "A", "B", "B", "B", "B", "C", "C", "C", "D", "D", "D", "E", "E", "F", "F", "G", "H", "I", "J"
    ), 6),
    nutrient = rep(c(rep(NA, 16), "Y", "Y", "Y", "Y", "Z", "Z", "Z", "Z", "Y", "Y", "Y", "Z", "Z", "Z", "Y", "Y", "Z", "Z", "Y", "Z", "Y", "Z"), 6),
    p_g = rep(NA, 38 * 6),
    p_e = rep(NA, 38 * 6)
  )
  
  env_or_x <- c(220, 230, 700, 705, 710, 750)
  env_or_y <- c(660, 610, 450, 430, 510, 550)
  
  # genetic predispositions:
  x_g <- sample(1:1000, 38)
  y_g <- sample(1:1000, 38)
  
  # ecological availability:
  x_e <- sample(1:1000, 38)
  y_e <- sample(1:1000, 38)
  
  for(behav in 1:38){
    output[output$behaviour == behav,]$p_g <- 1 - rescale(sqrt((x_g[behav] - env_or_x)^2 + (y_g[behav] - env_or_y)^2), to = c(1 - alpha_g, alpha_g))
    
    if(behav > 16){
      output[output$behaviour == behav,]$p_e <- 1 - rescale(sqrt((x_e[behav] - env_or_x)^2 + (y_e[behav] - env_or_y)^2), to = c(1 - alpha_e, alpha_e))
    }
  }
  # return the tibble:
  output
}

update_demography <- function(pop) {
  pop[, 39] <- pop[, 39] + 1
  pop[pop[, 39] >= 720, ] <- 0
  old <- which(pop[, 39] > 300)
  dead <- sample(c(TRUE, FALSE), length(old), prob = c(.01, .99), replace = TRUE)
  pop[old[dead], ] <- 0
  pop
}

update_social_behaviours <- function(pop, test_world, sd_peering) {
  N <- dim(pop)[1]
  state <- ((rowSums(pop[, 1:4]) >= 1) + (rowSums(pop[, 5:8]) >= 1) + (rowSums(pop[, 9:12]) >= 1) + (rowSums(pop[, 13:16]) >= 1)) / 4
  p_state <- runif(N) < rnorm(N, mean = 1 - state, sd = 0.05)
  p_peering <- rnorm(16, mean = colSums(pop[, 1:16]), sd = sd_peering)
  p_peering[p_peering < 0] <- 0
  innovation_i <- sample(1:16, N, prob = p_peering, replace = TRUE)
  p_innovate <- runif(N) < test_world$p_g[innovation_i]  * p_state
  for (i in (1:N)[p_innovate]) {
    pop[i, innovation_i[i]] <- 1
  }
  pop
}  

update_food_behaviours <- function(pop, test_world, sd_peering) {
  N <- dim(pop)[1]
  nut_y <- (rowSums(pop[, 17:20])>=1) + (rowSums(pop[, 25:27])>=1) + (rowSums(pop[, 31:32])>=1) + pop[, 35] + pop[, 37]
  nut_z <- (rowSums(pop[, 21:24])>=1) + (rowSums(pop[, 28:30])>=1) + (rowSums(pop[, 33:34])>=1) + pop[, 36] + pop[, 38]
  state <- (nut_y + nut_z - abs(nut_y - nut_z)) / 10 
  p_state <- runif(N) < rnorm(N, mean = 1 - state, sd = 0.05)
  p_peering <- rnorm(22, mean = colSums(pop[, 17:38]), sd = sd_peering)
  p_peering[p_peering < 0] <- 0
  innovation_i <- sample(17:38, N, prob = p_peering, replace = TRUE)
  p_innovate <- runif(N) < test_world$p_g[innovation_i] * test_world$p_e[innovation_i] * p_state
  for (i in (1:N)[p_innovate]) {
    pop[i, innovation_i[i]] <- 1
  }
  pop
}
```

However, we need to do some modifications to the main function `mockup_oranzees()`, as we want now, as output, the final composition of the population, and not the frequency of the traits:

```{r}
mockup_oranzees_pop <- function(t_max, alpha_g, alpha_e, sd_peering) {
  N <- 100
  oranzees_world <- set_oranzees_world(alpha_g, alpha_e)
  test_world <- oranzees_world %>%
    filter(population == "Uossob")
  pop <- matrix(c(rep(0, 38 * N), sample(1:300, N, replace = TRUE)), nrow = N, byrow = FALSE)
  # start simulation here:
  for (t in 1:t_max) {
    pop <- update_demography(pop)
    pop <- update_social_behaviours(pop, test_world, sd_peering)
    pop <- update_food_behaviours(pop, test_world, sd_peering)
  }
  output <- pop
}
```

The new function has less parameters and less complications, since it only runs a single test, and it gives as output all the population. We can run it, starting with the 'usual' parameter values:

```{r}
my_test <- mockup_oranzees_pop(t_max = 6000, alpha_g = 0.7, alpha_e = 0.9, sd_peering = 1)
```

We can write `head(my_test)` (remember it is a matrix), to see the first lines of the output:

```{r}
head(my_test)
```

Each of the first six lines shows an oranzee. A first aspect is, as said before, the distribution of traits among the population. 




***

The documentation continues in [05-oranzees.html](05-oranzees.html)

