---
title: "Oranzees model (06)"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

This R markdown file documents the development of the Oranzees model (it continues from [05-oranzees.html](56-oranzees.html)). The code-only version to run the model is in the file [main.R](main.R) in the same repository.

### Consolidating main.R

At this point, we have enough material to begin to run some extensive tests of the simulations. Before starting, however, it is useful to consolidate the code present in [main.R](main.R). We will reuse the functions that we have already written, with few small differences, i.e. slight improvments in the implementation (e.g. we wrote all the plotting routines as functions, etc.), or change of names

We can call all the functions with a direct call to [main.R](main.R):

```{r message=FALSE}
source('main.R')
```

Here a quick description of the **MAIN FUNCTIONS** and their associated **PLOTTING FUNCTIONS** in [main.R](main.R) (**COMMON FUNCTIONS** have been described at lenght in the previous documents):

#### `test_oranzees_1()`

`test_oranzees_1()` runs the simulation only on one population, and produces a richer output, consisting in the frequencies of all behaviours at each time step when `n_run=1`, and in the final frequencie of all behaviours when `n_run>1`. We can test that everything works by running and plotting the first case (single run, using `plot_one_run()`):

```{r message=FALSE, warning=FALSE}
my_test <- test_oranzees_1(t_max = 6000, alpha_g = 0.1, alpha_e = 0.9, init_world = TRUE, n_run = 1)
plot_one_run(my_test)
```

And in the second case (multiple runs, `using plot_multiple_runs()`):

```{r message=FALSE, warning=FALSE}
my_test <- test_oranzees_1(t_max = 6000, alpha_g = 0.1, alpha_e = 0.9, init_world = TRUE, n_run = 5)
plot_multiple_runs(my_test)
```

#### `test_oranzees_2()`

`test_oranzees_2()` also runs the simulation on one population, but gives as output the behavioural codes as described in Whiten et al., 1999. Here an example, using the associated plotting function `plot_codes_distribution()`:

```{r message=FALSE, warning=FALSE}
my_test <- test_oranzees_2(t_max = 6000, alpha_g = 0.1, alpha_e = 0.9, init_world = TRUE, n_run = 5)
plot_codes_distribution(my_test)
```

#### `run_oranzees()`

Finally, `run_oranzees()` runs the simulation on the six oranzees populations, and produces as output the code for each behaviour in each population. While it works with several run, the only possible visualisation, for the time being, concerns a single run, as illustrated below. Notice the associated plotting function `plot_oranzees()` has a further parameter `social`. When `social=TRUE`, social behaviours are visualised, when `social=FALSE`, it visualises food-related behaviours.

```{r message=FALSE, warning=FALSE}
my_test <- run_oranzees(t_max = 6000, alpha_g = 0.1, alpha_e = 0.9, init_world = TRUE, n_run = 1)
```

```{r message = FALSE, fig.height = 14}
plot_oranzees(my_test, social = TRUE)
```

```{r message = FALSE, fig.height = 14}
plot_oranzees(my_test, social = FALSE)
```

### Estimating oranzees population sizes

So far, we tested `run_oranzees()` with arbitrary values for population sizes (`N <- c(50, 100, 150, 50, 100, 150)`). Before starting the actual complete tests, we need to estabilish population sizes that are consistent with the data in Whiten et al., 1999.

According to the paper, the six groups used in Figure 1 are Bossou, Kibale, Budongo, Taï, Gombe, and Mahale (M-group). No information on population sizes are, however, reported there. 

Lind and Lindenfors, 2010, tested the correlation between the number of cultural traits in the seven communities of Whiten et al. (including also Mahale K-group, in which we are not interested) and their population size. They report that data from community sizes were from Wrangham, 2000. We use the same data (from table 21.1 in Wrangham, 2000), rounding them to the highest integer:

Group     |  Population size  
--------- |----
Bossou    | 20
Kibale    | 42
Budongo   | 49
Taï       | 76  
Gombe     | 50
Mahale(M) | 95

According to the order in which our oranzees community are listed in the code ("Uossob", "Iat Forest", "Ebmog", "Elaham", "Elabik", "Ognodub"), this translated in:

```{r}
N <- c(20, 76, 50, 95, 42, 49)
```

which we can now include in the function `run_oranzees()` in [main.R](main.R).

***


