---
title: "Oranzees model"
author: "Alberto Acerbi"
date: "13/05/2019"
output: github_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This Rmarkdown file documents the development of the Oranzees model. The code-only version to run the model is in the file *main.R* in the same repository.

### Spatial set-up

We first need to set-up a simulated environment where six populations of oranzees live at relative distances that are approximately similar to the six populations of chimpanzees described in Whiten et al., 1999 

We use the R package [imager](https://dahtah.github.io/imager/imager.html) to process a capture from Figure 1 in Whiten et al., 1999.

```{r message = FALSE}
library(imager)

chimp_env <- load.image("material/Whiten_1999_cap.png")
plot(chimp_env)
```

We manually process the image to have only the six points to which we are interested, and we also reverse the y-axis (in image processing starting from zero and go negative is standard, but this is not useful for us):

```{r}

chimp_env <- load.image( "material/Whiten_1999_cap2.png")
plot(chimp_env, axes=F)
axis(1)
axis(2, at = seq(-100, 400, 100), labels = seq(500, 0, -100))
grid()
box()

```

Here are the approximate locations of the oranzees populations:

```{r}

env_or_x <- c(420, 430, 900, 905, 910, 950)
env_or_y <- c(360, 310, 150, 130, 210, 250)

plot(env_or_x, env_or_y, xlim = c(0, 1200), ylim = c(0, 500), pch = 21, bg = "red")
grid()

```

For convenience, we put these locations approximately in the centre of a 1000,1000 squared environment (what is important are the relative distances, not their absolute values):

```{r}

env_or_x <- env_or_x - 200
env_or_y <- env_or_y + 300

par(pty = "s")
plot(env_or_x, env_or_y, xlim = c(0, 1000), ylim = c(0, 1000), pch = 21, bg = "red")
grid()

```

### Genetic and environmental gradients