---
title: "Oranzees model"
author: "Alberto Acerbi"
output: html_document

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


This R markdown file documents the development of the Oranzees model. The code-only version to run the model is in the file *main.R* in the same repository.

## Spatial set-up

We first need to set-up a simulated environment where six populations of oranzees live at relative distances that are approximately similar to the six populations of chimpanzees described in Whiten et al., 1999. 

We use the R package [`imager`](https://dahtah.github.io/imager/imager.html) to process a capture from Figure 1 in Whiten et al., 1999.

```{r message = FALSE}
library(imager)
chimp_env <- load.image("material/Whiten_1999_cap.png")
plot(chimp_env)
box()
```

We manually process the image to have only the six points to which we are interested, and we also reverse the y-axis (in image processing starting from zero and go negative is standard, but this is not useful for us):

```{r}
chimp_env <- load.image("material/Whiten_1999_cap2.png")
plot(chimp_env, axes = F)
axis(1)
axis(2, at = seq(-100, 400, 100), labels = seq(500, 0, -100))
grid()
box()
```

Here are the approximate locations of the oranzees populations:

```{r}
env_or_x <- c(420, 430, 900, 905, 910, 950)
env_or_y <- c(360, 310, 150, 130, 210, 250)

plot(env_or_x, env_or_y, xlim = c(0, 1200), ylim = c(0, 500), pch = 21, bg = "red")
grid()
```

For convenience, we put these locations approximately in the centre of a 1000,1000 squared environment (what is important are the relative distances, not their absolute values):

```{r}
env_or_x <- env_or_x - 200
env_or_y <- env_or_y + 300

par(pty = "s")
plot(env_or_x, env_or_y, xlim = c(0, 1000), ylim = c(0, 1000), pch = 21, bg = "red")
grid()
```

## The behaviours

Following again Whiten et al., 1999, we consider 38 possible behaviours. We distinguish among 16 'social' and 22 'food-related' behaviours.


### Genetic propensity

We first implement the idea that the probability to acquire any of these behaviours depends (among other things) by what we call 'genetic propensity', and we model as a probability $p_g(0,1)$.

These probability is assigned randomly and independently to each behaviours. To assign it to the different populations we use a geographical gradient. For each behaviour, a random point is chosen in the Cartesian space and its distance to each population is calculated. This distance is then normalised between 0 and 1000. The actual probabilities are then calculated as $1-d_{norm}$.

Here an example:

```{r}
# generate the random point for the genetic propensity:
x_g <- sample(1:1000, 1)
y_g <- sample(1:1000, 1)

# calcualte the distance from the six populations:
d_g <- c()
for (i in 1:6) {
  d_g[i] <- sqrt((x_g - env_or_x[i])^2 + (y_g - env_or_y[i])^2)
}
```

Now we can see the random point we extracted:

```{r echo=FALSE}
c(x_g, y_g)
```

And the distances calculated:

```{r echo=FALSE}
d_g
```

We normalise the distance by 1000, and we then find the final probability subtracting from 1 (values that are below 0, i.e. population that were at more than 1000 units distance, are considered 0):

```{r}
d_g_norm <- d_g / 1000
p_g <- 1 - d_g_norm
p_g[p_g < 0] <- 0
```

Here are the final values for $p_g$:

```{r echo=FALSE}
p_g
```

An intuitive way to see this graphically is to use a gradient. The code is slightly more complicated, but we just need to run it once to visualise the effect. We first calculate the values of $p_g$ for all the points in the environment:

```{r}
gradient_g <- matrix(nrow = 1000, ncol = 1000)

for (i in 1:1000) {
  for (j in 1:1000) {
    gradient_g[i, j] <- sqrt((x_g - i)^2 + (y_g - j)^2)
  }
}
gradient_g <- 1 - gradient_g / 1000
gradient_g[gradient_g < 0] <- 0
```

And we then plot it, using the R package [`viridis`](https://cran.r-project.org/web/packages/viridis/vignettes/intro-to-viridis.html), for the colour scale.

```{r message = FALSE}
library(viridis)
par(pty = "s")
image(1:1000, 1:1000, gradient_g, col = viridis(10), xlab = "env_or_x", ylab = "env_or_y")
contour(1:1000, 1:1000, gradient_g, add = TRUE)
points(env_or_x, env_or_y, pch = 21, bg = "red")
```

We can check that the $p_g$ values of the six different populations look like the ones we showed above, just before working on the visualisation (the order of the populations is from left to right on the x-axis). Remember, this means that population with higher values will be more genetically predisposed to acquire one of the 38 behaviours.

### Ecological availability

We define, together with genetic propensity, a similar variable $p_e(0,1)$, standing for ecological availability. Whereas the likelihood of acquiring all beahviours depends on $g_e$, ecological availability has influence only on food-related behaviours. Ecological availability can be thought as the probability of finding a resource, or its nutritional values, in one of the site.

ecological availability is implemented as genetic propensity, with the only difference that the distances are normalised by 500:

```{r}
x_e <- sample(1:1000, 1)
y_e <- sample(1:1000, 1)

d_e <- c()
for (i in 1:6) {
  d_e[i] <- sqrt((x_e - env_or_x[i])^2 + (y_e - env_or_y[i])^2)
}

d_e_norm <- d_e / 500
p_e <- 1 - d_e_norm
p_e[p_e < 0] <- 0
```

As before, here are the final values for $p_e$:

```{r echo=FALSE}
p_e
```

And we can visualise them, using the same technique:

```{r}
gradient_e <- matrix(nrow = 1000, ncol = 1000)

for (i in 1:1000) {
  for (j in 1:1000) {
    gradient_e[i, j] <- sqrt((x_e - i)^2 + (y_e - j)^2)
  }
}
gradient_e <- 1 - gradient_e / 500
gradient_e[gradient_e < 0] <- 0

par(pty = "s")
image(1:1000, 1:1000, gradient_e, col = magma(10), xlab = "env_or_x", ylab = "env_or_y")
contour(1:1000, 1:1000, gradient_e, add = TRUE)
points(env_or_x, env_or_y, pch = 21, bg = "red")
```

Again, we can check that the values in the plot correspond to the ones calculated above. The difference between the plot for genetic propensity and the plot for ecological availability illustrates the difference in the normalisation (1000 for the former and 500 for the latter). In the ecological availability plot, depending where the random point has been placed, it is more likely that some areas of the environment will be with $p_e=0$, including possibly some of our six populations. This reflects the fact that, for example, some raw materials can be available in a population, but not in another. The likelihood of $p_g=0$ is instead lower, reflecting that genetic constraints are less strong.


### Categories of behaviours

We model the idea that different behaviours can serve the same goal, so that if behaviour $1$ and $2$ serve the same goal, an individual performing $1$ will be unlikely to also perform $2$. 

In the case of social behaviours, we assume 4 categories of behaviours, each with 4 possible different behaviours, that serve the same goal, as illustrated here (notice the categories' names are purely indicative):

Category  |    |    |    | 
--------- |----|----|----|----
play      | 1  | 2  | 3  | 4
display   | 5  | 6  | 7  | 8
groom     | 9  | 10 | 11 | 12
courtship | 13 | 14 | 15 | 16

We do the same for food-related behaviours, with the difference that, for each 'goal' there is a variable number of behaviours, as such:

Category |    |    |    | 
---------|----|----|----|----
A        | 17 | 18 | 19 | 20
B        | 21 | 22 | 23 | 24
C        | 25 | 26 | 27 | 
D        | 28 | 29 | 30 | 
E        | 31 | 32 |    |
F        | 33 | 34 |    |
G        | 35 |    |    |
H        | 36 |    |    |
I        | 37 |    |    |
J        | 38 |    |    |

In addition, the categories of food-related behaviours are associated to two different 'nutrients'. The idea is that individuals need to balance their nutritional intake, so that their optimal diet consist in a roughly equal number of foodstuff for one and the other nutrient. Our final classification of food-related behaviours is:

Category |    |    |    |    | Nutrient
---------|----|----|----|----|---------
A        | 17 | 18 | 19 | 20 |    Y
B        | 21 | 22 | 23 | 24 |    Z
C        | 25 | 26 | 27 |    |    Y
D        | 28 | 29 | 30 |    |    Z
E        | 31 | 32 |    |    |    Y
F        | 33 | 34 |    |    |    Z
G        | 35 |    |    |    |    Y
H        | 36 |    |    |    |    Z
I        | 37 |    |    |    |    Y
J        | 38 |    |    |    |    Z 

***

We can now wrap everything up in a function `set_oranzees_environment()` that can be used independently from what we done so far, and that we will use in the code of the actual model (*main.R*):

```{r message=FALSE}
library(tidyverse)

set_oranzees_environment <- function() {
  output <- tibble(
    population = as.factor(rep(1:6, each = 38)),
    behaviour = as.factor(rep(1:38, 6)),
    type = rep(c(rep("social", 16), rep("food-related", 22)), 6),
    category = rep(c(
      rep("play", 4), rep("display", 4), rep("groom", 4), rep("courthsip", 4),
      "A", "A", "A", "A", "B", "B", "B", "B", "C", "C", "C", "D", "D", "D", "E", "E", "F", "F", "G", "H", "I", "J"
    ), 6),
    nutrient = rep(c(rep(NA, 16), "Y", "Y", "Y", "Y", "Z", "Z", "Z", "Z", "Y", "Y", "Y", "Z", "Z", "Z", "Y", "Y", "Z", "Z", "Y", "Z", "Y", "Z"), 6),
    p_g = rep(NA, 38 * 6),
    p_e = rep(NA, 38 * 6)
  )

  env_or_x <- c(220, 230, 700, 705, 710, 750)
  env_or_y <- c(660, 610, 450, 430, 510, 550)

  # genetic predispositions:
  x_g <- sample(1:1000, 38)
  y_g <- sample(1:1000, 38)

  # ecological availability:
  x_e <- sample(1:1000, 38)
  y_e <- sample(1:1000, 38)

  for (oranzees_pop in 1:6) {
    for (behav in 1:38) {
      output[output$population == oranzees_pop & output$behaviour == behav, ]$p_g <- 1 - sqrt((x_g[behav] - env_or_x[oranzees_pop])^2 + (y_g[behav] - env_or_y[oranzees_pop])^2) / 1000

      if (behav > 16) { # only for food-related behaviours:
        output[output$population == oranzees_pop & output$behaviour == behav, ]$p_e <- 1 - sqrt((x_e[behav] - env_or_x[oranzees_pop])^2 + (y_e[behav] - env_or_y[oranzees_pop])^2) / 500
      }
    }
  }

  # replace negative values with 0s:
  if (dim(output[output$p_g < 0, ])[1] > 0) {
    output[output$p_g < 0, ]$p_g <- 0
  }
  if (dim(output[output$p_e < 0 & !is.na(output$p_e), ])[1] > 0) {
    output[output$p_e < 0 & !is.na(output$p_e), ]$p_e <- 0
  }

  # return the tibble:
  output
}
```

This code does what we already did, but for all the 38 behaviours, and store everything in a 'tibble' structure, provided by the library `tidyverse`. In the main code, we just need to run the function to start the set-up of a simulation:

```{r}
oranzees_environment <- set_oranzees_environment()
```

We can explore the output simply writing the name of the data structure:

```{r}
oranzees_environment
```

Each line of the tibble *oranzees_environment* records the information for each behaviour, for each population. Since there are 38 behaviours and 6 populations, the tibble has 228 lines. You can notice that the values of $nutrient$ and $p_e$ for behaviours of the $social$ category are NA, that is "not available", which is what we described above.

There are a couple of checks that we can do to be sure that the function works as desired. First, we would expect, on average, genetic propensities being higher than ecological availabilities, and that it is more likely that some of the latter will be equal to zero. We can use the `facet` functionality in `ggplot` to split the data for the six populations. Here are the results for $p_g$:

```{r}
ggplot(data = oranzees_environment) +
  geom_point(aes(x = behaviour, y = p_g)) +
  facet_wrap(~population) +
  theme_bw() +
  labs(y = "Genetic propensity") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

We can do the same plot for $p_e$:

```{r warning=F}
ggplot(data = oranzees_environment) +
  geom_point(aes(x = behaviour, y = p_e)) +
  facet_wrap(~population) +
  theme_bw() +
  labs(y = "Ecological availability") +
  theme(axis.text.x = element_blank(), axis.ticks.x = element_blank())
```

As we wanted, values of $p_e$ are lower, and a few of them are zeros (also notice that the first 16 values for each population, on the left side of the plots, are not present as they are social behaviours). 

A second check exploits the fact that both $p_e$ and $p_g$ are assigned to the six populations following a geographical gradient. We would expect, thus, that populations that are geographically close to each other are also more similar genetically and they live in a similar environment. In other words, we expect for example populations 1 and 2 to be more similar one to each other than to populations 3 (or 4, 5, and 6).

We can check this with some quick scatterplots. Here the correlation between the genetic propensities of populations 1 and 2 (black dots) and 1 and 3 (red dots) for all 38 behaviours:

```{r}
par(pty = "s")
plot(oranzees_environment[oranzees_environment$population == 1, ]$p_g, oranzees_environment[oranzees_environment$population == 2, ]$p_g,
  pch = 19,
  xlab = "Population 1", ylab = "Population check", xlim = c(0, 1), ylim = c(0, 1)
)
points(oranzees_environment[oranzees_environment$population == 1, ]$p_g, oranzees_environment[oranzees_environment$population == 3, ]$p_g, pch = 19, col = "red")
grid()
```

And the same for the the ecological availabilities of the 22 food-related behaviours:

```{r}
par(pty = "s")
plot(oranzees_environment[oranzees_environment$population == 1, ]$p_e, oranzees_environment[oranzees_environment$population == 2, ]$p_e,
  pch = 19,
  xlab = "Population 1", ylab = "Population check", xlim = c(0, 1), ylim = c(0, 1)
)
points(oranzees_environment[oranzees_environment$population == 1, ]$p_e, oranzees_environment[oranzees_environment$population == 3, ]$p_e, pch = 19, col = "red")
grid()
```



