---
title: "Simulations for manuscript"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All codes are in [test.R](test.R)

```{r message=FALSE}
source("main.R")
```

### Figure 1 

Data from AWS

### Figure 2

How all patterns ("A", "B", "C", "D") corresponds to the ones found in Whiten et al., 1999, in a condition for which we have a good match for "cultural" behaviours (pattern "D"). As above the code is not run, but the data are loaded. 

```{r eval=FALSE}
tic()
test <- run_oranzees(t_max = 6000, alpha_e = 0.8, alpha_g = 0.2, S = 1, init_world = TRUE, n_run = 20)
toc()
write_csv(test, "sims_for_ms/data/figure_2.csv")
```

```{r message=FALSE}
whiten <- tibble(pattern = 1:4, 
                 n = c(7, 16, 3, 38), 
                 data = rep("Whiten", 4)) 
figure_2 <- analyse_patterns(read_csv("sims_for_ms/data/figure_2.csv"))
ggplot(data = figure_2, aes(x = pattern, y = n)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width=0.05, alpha=0.5) +
    theme_bw() +
    geom_line(data = whiten, colour="red") +
    geom_point(data = whiten, colour="red") +
    ggsave("sims_for_ms/figures/figure_2.pdf")
```

### Figure 3

The effect of $S$, in three conditions: (a) no genetic differences and intermediate ecological differences (high-left angle of figure 1, produces too few "cultural" traits), (b) good match with Whiten, and (c) intermediate genetic differnces and high ecological differences (low-right angle of figure 1, produces too many "cultural" traits).

Below the example code for condition (a).

```{r eval=FALSE}
tic()
S = seq(0, 1, by = 0.1)
output <- tibble(S = rep(S, each = 10), n = NA)
 for(s in S){
  test <- run_oranzees(t_max = 6000, alpha_e = 0.5, alpha_g = 0, S = s, init_world = TRUE, n_run = 10)
  out <- analyse_patterns(test)
  output[output$S == s, ]$n <- out$n[31:40]
  print(s)
}
toc()
write_csv(test, "sims_for_ms/data/figure_3_a.csv")
```

```{r message=FALSE, fig.width=10}
a <- read_csv("sims_for_ms/data/figure_3_a.csv")
b <- read_csv("sims_for_ms/data/figure_3_b.csv")
c <- read_csv("sims_for_ms/data/figure_3_c.csv")

figure_3 <- tibble(S = as.factor(c(a$S, b$S, c$S)), n = c(a$n, b$n, c$n), 
       alpha_g = rep(c(0, 0.2, 0.5), each = 110), 
        alpha_e = rep(c(0.5, 0.8, 1), each=110)) 

figure_3_mean <- figure_3 %>%
  group_by(S, alpha_g, alpha_e) %>%
  summarise(cultural = mean(n))

figure_3 <- figure_3 %>%
  full_join(figure_3_mean)

ggplot(data = figure_3, aes(y=n, x=S)) +
  geom_boxplot(outlier.shape = NA, aes(fill=cultural)) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1), aes(fill=cultural)) +
  facet_wrap(alpha_e ~ alpha_g, labeller=label_both) +
  scale_fill_gradientn(colors=c("blue","blue","white","red"), name="Number of\n'cultural' traits") +
  theme_bw() +
  ggsave("sims_for_ms/figures/figure_3.pdf", width = 12)
```