---
title: "Simulations for manuscript"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

All codes are also in [test.R](test.R). We need to load the usual functions in [main.R](main.R).

```{r message=FALSE}
source("main.R")
```

### Figure 1 

We first check the number of "cultural" behaviours for all the *interesting* combinations of the parameter $\alpha_e$ and $\alpha_g$, i.e. moderate to high environmental variability ($\alpha_e=(0.5,1)$) and zero to moderate genetic difference ($\alpha_g=(0,0.5)$). We run 20 simulations for each combination (20*36 runs, code is not run here).

```{r eval=FALSE}
alpha_e_test = seq(0.5, 1, by = 0.1)
alpha_g_test = seq(0, 0.5, by = 0.1)

output <- tibble(alpha_e = rep(alpha_e_test, each = 120), 
                 alpha_g = rep(rep(alpha_g_test, each = 20), 6),
                 n = NA)

for(e in alpha_e_test){
  for(g in alpha_g_test) {
    test <- run_oranzees(t_max = 6000, alpha_e = e, alpha_g = g, S = 1, init_world = TRUE, n_run = 20)
    out <- analyse_patterns(test)
    output[output$alpha_e == e & output$alpha_g == g, ]$n <- out$n[61:80]
    print(e)
  }
}
write_csv(output, "sims_for_ms/data/figure_1.csv")
```

##### Visualisation 1 (matrix)

```{r  message=FALSE}
figure_1 <- read_csv("sims_for_ms/data/figure_1.csv") %>%
  group_by(alpha_g, alpha_e) %>%
  summarise( cultural = mean(n))
ggplot(data=figure_1) +
  geom_raster(aes(x = alpha_g, y = alpha_e, fill = cultural)) +
  scale_fill_gradientn(colors=c("red","white","blue")) +
  geom_text(aes(x = alpha_g, y = alpha_e, label = cultural))+
  theme_bw()
```

##### Visualisation 2 (histogram)

```{r message=FALSE, fig.width=10, fig.height=7}
figure_1_mean <- read_csv("sims_for_ms/data/figure_1.csv") %>%
  group_by(alpha_g, alpha_e) %>%
  summarise( cultural = mean(n))
figure_1 <- read_csv("sims_for_ms/data/figure_1.csv") %>%
  full_join(figure_1_mean)
ggplot(data = figure_1) + 
  geom_histogram(aes(x = n, fill=cultural), color="black", bins=10) +
  geom_vline(xintercept = 39, linetype="dashed") +
  scale_fill_gradientn(colors=c("blue","white","red"), name="Number of\n'cultural' traits") +
  facet_grid(alpha_e ~ alpha_g, labeller=label_both) +
  theme_bw() +
  labs(y = "Number of runs", x = "Number of 'cultural' traits") +
  ggsave("sims_for_ms/figures/figure_1.pdf", height=7, width = 10)

```

##### Visualisation 3 (density)

```{r message=FALSE, fig.width=10, fig.height=7}
figure_1_mean <- read_csv("sims_for_ms/data/figure_1.csv") %>%
  group_by(alpha_g, alpha_e) %>%
  summarise( cultural = mean(n))
figure_1 <- read_csv("sims_for_ms/data/figure_1.csv") %>%
  full_join(figure_1_mean)
ggplot(data = figure_1) + 
  geom_density(aes(x = n, fill=cultural)) +
  geom_vline(xintercept = 39, linetype="dashed") +
  scale_fill_gradientn(colors=c("blue","white","red"), name="Number of\n'cultural' traits") +
  facet_grid(alpha_e ~ alpha_g, labeller=label_both) +
  theme_bw()
```

***

### Figure 2

How all patterns ("A", "B", "C", "D") corresponds to the ones found in Whiten et al., 1999, in a condition for which we have a good match for "cultural" behaviours (pattern "D"). As above the code is not run, but the data are loaded. 

```{r eval=FALSE}
tic()
test <- run_oranzees(t_max = 6000, alpha_e = 0.8, alpha_g = 0.2, S = 1, init_world = TRUE, n_run = 20)
toc()
write_csv(test, "sims_for_ms/data/figure_2.csv")
```

```{r message=FALSE}
whiten <- tibble(pattern = 1:4, 
                 n = c(7, 16, 3, 38), 
                 data = rep("Whiten", 4)) 
figure_2 <- analyse_patterns(read_csv("sims_for_ms/data/figure_2.csv"))
ggplot(data = figure_2, aes(x = pattern, y = n)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width = 0.05, alpha = 0.5) +
    theme_bw() +
    geom_line(data = whiten, colour="red") +
    geom_point(data = whiten, colour="red") +
    labs(y = "Number of traits") +
    ggsave("sims_for_ms/figures/figure_2.pdf")
```

We also use the same data to calcualte the correlation between population size and number of traits (habitual + customary + present).

```{r message=FALSE, fig.width=5}
data <- read_csv("sims_for_ms/data/figure_2.csv")
pop_sizes <- c(20, 76, 50, 95, 42, 49)
list_pop <- c("Uossob", "Iat Forest", "Ebmog", "Elaham", "Elabik", "Ognodub")

customary <- c()
habitual <- c()
present <- c()

for( r in 1:20){
  for(pop in list_pop){
    customary <- c(customary, dim(data[data$run==r & data$population==pop & data$code=="customary",])[1])
    habitual <- c(habitual, dim(data[data$run==r & data$population==pop & data$code=="habitual",])[1])
    present <- c(present, dim(data[data$run==r & data$population==pop & data$code=="present",])[1])
  }
}
figure_2_extra <- tibble(behaviours = customary + present + habitual, 
              pop_size = rep(pop_sizes, 20))
ggplot(data = figure_2_extra, aes(x=pop_size, y=behaviours)) +
  geom_jitter(width = 0.05, alpha = 0.5) +
  theme_bw() +
  stat_smooth(method = lm)  +
  labs(x="Population size", y="Number of traits")+
  ggsave("sims_for_ms/figures/figure_2_extra.pdf", width = 5)
```

***

### Figure 3

The effect of $S$, in three conditions: (a) no genetic differences and intermediate ecological differences (high-left angle of figure 1, produces too few "cultural" traits), (b) good match with Whiten, and (c) intermediate genetic differnces and high ecological differences (low-right angle of figure 1, produces too many "cultural" traits).

Below the example code for condition (a).

```{r eval=FALSE}
tic()
S = seq(0, 1, by = 0.1)
output <- tibble(S = rep(S, each = 10), n = NA)
 for(s in S){
  test <- run_oranzees(t_max = 6000, alpha_e = 0.5, alpha_g = 0, S = s, init_world = TRUE, n_run = 10)
  out <- analyse_patterns(test)
  output[output$S == s, ]$n <- out$n[31:40]
  print(s)
}
toc()
write_csv(test, "sims_for_ms/data/figure_3_a.csv")
```

```{r message=FALSE, fig.width=10}
a <- read_csv("sims_for_ms/data/figure_3_a.csv")
b <- read_csv("sims_for_ms/data/figure_3_b.csv")
c <- read_csv("sims_for_ms/data/figure_3_c.csv")

figure_3 <- tibble(S = as.factor(c(a$S, b$S, c$S)), n = c(a$n, b$n, c$n), 
       alpha_g = rep(c(0, 0.2, 0.5), each = 110), 
        alpha_e = rep(c(0.5, 0.8, 1), each = 110)) 

figure_3_mean <- figure_3 %>%
  group_by(S, alpha_g, alpha_e) %>%
  summarise(cultural = mean(n))

figure_3 <- figure_3 %>%
  full_join(figure_3_mean)

ggplot(data = figure_3, aes(y = n, x = S)) +
  geom_boxplot(outlier.shape = NA, aes(fill = cultural)) +
  geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1), aes(fill=cultural)) +
  facet_wrap(alpha_e ~ alpha_g, labeller = label_both) +
  scale_fill_gradientn(colors = c("blue", "blue", "white","red"), name = "Number of\n'cultural' traits") +
  theme_bw() +
  labs(x = "S (probability of socially-mediated innovation)", y = "Number of 'cultural' traits") +
  ggsave("sims_for_ms/figures/figure_3.pdf", width = 12)
```


