---
title: "Oranzees first tests results"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Codes for behavioural patterns

Whiten et al., 1999 extract the above patterns from their data:

* *A*: patterns absent at no site.

* *B*: patterns not achieving habitual frequencies at any site.

* *C*: patterns for which any absence can be explained by local ecological factors.

* *D*: patterns customary or habitual at some sites yet absent at others, with no ecological explanation, i.e. the "cultural" behaviours.

We thus need to write a further function, `analyse_patterns()`, in [test.R](test.R), that takes as input the output of `run_oranzees()`, and produces the number of behaviours for each pattern in each run:

```{r}
analyse_patterns <- function(my_test){
  output <- tibble(pattern = rep(c("A", "B", "C", "D"), each = max(my_test$run)), 
                   n = rep(NA, max(my_test$run) * 4))
  process <- my_test %>%
    group_by(behaviour, run) %>%
    summarise(A = !("absent" %in% code) && !("ecological explanation" %in% code), 
              B = !("habitual" %in% code) && !("customary" %in% code),
              C = !("absent" %in% code) && ("ecological explanation" %in% code)) %>%
    ungroup() %>%
    mutate (D = !(A | B | C)) %>%
    group_by(run)
  
  temp <- count(process, A) %>%
    filter(A == TRUE)
  out <- temp$n
  if(length(temp$n) < max(my_test$run)){
    out <- c(temp$n, rep(0, max(my_test$run)-length(temp$n)))
  }
  output[output$pattern=="A",]$n <- out  
  
  temp <- count(process, B) %>%
    filter(B == TRUE)
  out <- temp$n
  if(length(temp$n) < max(my_test$run)){
    out <- c(temp$n, rep(0, max(my_test$run)-length(temp$n)))
  }
  output[output$pattern=="B",]$n <- out  
  
  temp <- count(process, C) %>%
    filter(C == TRUE)
  out <- temp$n
  if(length(temp$n) < max(my_test$run)){
    out <- c(temp$n, rep(0, max(my_test$run)-length(temp$n)))
  }
  output[output$pattern=="C",]$n <- out 
  
  temp <- count(process, D) %>%
    filter(D == TRUE)
  out <- temp$n
  if(length(temp$n) < max(my_test$run)){
    out <- c(temp$n, rep(0, max(my_test$run)-length(temp$n)))
  }
  output[output$pattern=="D",]$n <- out 
  
  output
}

```

### Running the tests

Before starting, we need all the functions present in [main.R](main.R), as well as the library `tictoc`, to have information on the time needed for the simulations to run.

```{r message=FALSE}
library(tictoc)
source("main.R")
```

We run 20 simulations for $t_\text{max}=6000$, reinitialising the world for each simulations, for the following combination of parameters:

$\alpha_e$ | $\alpha_g$ 
------|-------
0     | 0
0     | 0.5
0     | 1
0.5   | 0
0.5   | 0.5 
0.5   | 1
1     | 0
1     | 0.5
1     | 1

The code, for example for the first combination, is:

```{r eval=FALSE}
tic()
test <- run_oranzees(t_max = 6000, alpha_e = 0, alpha_g = 0, init_world = TRUE, n_run = 20)
toc()
write_csv(test, "output_test/test_0_0.csv")
```

We run seprately the simulations (see TEST 1 in [test.R](test.R)), and simply load here the results:

```{r message=FALSE}
test_0_0 <-analyse_patterns(read_csv("output_test/test_0_0.csv"))
test_0_05 <-analyse_patterns(read_csv("output_test/test_0_05.csv"))
test_0_1 <-analyse_patterns(read_csv("output_test/test_0_1.csv"))
test_05_0 <-analyse_patterns(read_csv("output_test/test_05_0.csv"))
test_05_05 <-analyse_patterns(read_csv("output_test/test_05_05.csv"))
test_05_1 <-analyse_patterns(read_csv("output_test/test_05_1.csv"))
test_1_0 <-analyse_patterns(read_csv("output_test/test_1_0.csv"))
test_1_05 <-analyse_patterns(read_csv("output_test/test_1_05.csv"))
test_1_1 <-analyse_patterns(read_csv("output_test/test_1_1.csv"))
```

We can now plot the number of traits considered "cultural", according to the patterns in Whiten et al., 1999 in each condition:

```{r}
tibble(alpha_e = rep(c("0", "0.5", "1"), each = 60), 
       alpha_g = rep(c("0", "0.5", "1","0", "0.5", "1","0", "0.5", "1"), each = 20),
      behaviours = c(test_0_0[test_0_0$pattern=="D",]$n, 
                     test_0_05[test_0_05$pattern=="D",]$n,
                     test_0_1[test_0_1$pattern=="D",]$n,
                     test_05_0[test_05_0$pattern=="D",]$n,
                     test_05_05[test_05_05$pattern=="D",]$n,
                     test_05_1[test_05_1$pattern=="D",]$n,
                     test_1_0[test_1_0$pattern=="D",]$n,
                     test_1_05[test_1_05$pattern=="D",]$n,
                     test_1_1[test_1_1$pattern=="D",]$n)) %>%
  ggplot(aes( x = alpha_g, y = behaviours, fill = alpha_e)) +
    geom_boxplot(outlier.shape = NA) +
    geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1)) +
  scale_fill_manual(values = c("#FFDB6D", "#F4EDCA", "#D16103")) +
    theme_bw() +
    labs(y = "Number of traits considered 'cultural'") 
```

Whiten et al., 1999 found 38 "cultural" behaviour out of a total of 65 considered. This is close, for these parameters, to the conditions with high ecological variability ($\alpha_e=1$) and no genetic differences ($\alpha_g=0$), or to the condition with no ecological variability ($\alpha_e=0$) and intermediate genetic differences ($\alpha_g=0.5$).  It is useful to look more in details to this conditions to have a clearer picture of how the results look like. 

For example, what about the other patterns? We can plot all the pattern proportions in the conditions we are interested, and compare them directly with the results of Whiten et al., 1999.

```{r}
whiten <- tibble(pattern = 1:4, 
                 n = c(7, 16, 3, 38), 
                 data = rep("Whiten", 4)) 
ggplot(data = test_1_0, aes(x = pattern, y = n)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width=0.05, alpha=0.5) +
    theme_bw() +
    geom_line(data = whiten, colour="red") +
    geom_point(data = whiten, colour="red") 
```

```{r}
whiten <- tibble(pattern = 1:4, 
                 n = c(7, 16, 3, 38), 
                 data = rep("Whiten", 4)) 
ggplot(data = test_0_05, aes(x = pattern, y = n)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width=0.05, alpha=0.5) +
    theme_bw() +
    geom_line(data = whiten, colour="red") +
    geom_point(data = whiten, colour="red") 
```

With thse encouraging results, we can run a more extensive test, checking only the proportion of "cultural" behaviours (pattern D), for a finer-grained parameter space. We focus on the more realistic parameters, i.e. relatively low genetic differences (i.e. $\alpha_g=<0.5$) and relatively high environmental variability (i.e. $alpha_e>=0.5$).

Here is the code (not run, see TEST 2 in [test.R](test.R)):

```{r eval=FALSE}
alpha_e_test = seq(0.5, 1, by = 0.1)
alpha_g_test = seq(0, 0.5, by = 0.1)

output <- tibble(alpha_e = rep(alpha_e_test, each = 30), 
                 alpha_g = rep(rep(alpha_g_test, each = 5), 6),
                 n = rep(NA, 180))

for(e in alpha_e_test){
  for(g in alpha_g_test) {
    test <- run_oranzees(t_max = 6000, alpha_e = e, alpha_g = g, init_world = TRUE, n_run = 5)
    out <- analyse_patterns(test)
    output[output$alpha_e == e & output$alpha_g == g, ]$n <- out$n[16:20]
    print(e)
  }
}
write_csv(output, "output_test/test_many.csv")
```

Now to plot:

```{r message=FALSE}
test_many <- read_csv("output_test/test_many.csv") %>%
  group_by(alpha_g, alpha_e) %>%
  summarise( cultural = mean(n))
ggplot(data=test_many) +
  geom_raster(aes(x = alpha_g, y = alpha_e, fill = cultural)) +
  scale_fill_gradientn(colors=c("red","white","blue")) +
  geom_text(aes(x = alpha_g, y = alpha_e, label = cultural))+
  theme_bw()
```

### The importance of "socially-mediated innovation"

Many combinations of values for $\alpha_g$ and $\alpha_e$ seem to produce results that are similar to what found in Whiten et al., 1999, including "realistic" combinations, with low values of $\alpha_g$ and high values of $\alpha_e$ (i.e. genetic differences are not particularly important, whereas ecological ones are). In the manuscript, we will present in details an analysis of the $\alpha_g$ \ $\alpha_e$ parameter space in this respect.

Another intersting question concerns the importance of "socially-mediated innovation": how important it is? In other words, do "cultural" traits emerge because of it, or because of the other constraints, such as ecology, physiology, genetic, etc.?  

To test this, we need to add a last parameter to our model - let's call it $S$ - which gives the probability that each innovation is socially-mediated (i.e. its probability is a function of the frequency of the behaviour in the population), or random. In order to this, we need to add three lines to the functions updating the behaviour. Here is what we do in `update_social_behaviours()`:

```{r eval=FALSE}
  innovation_random <- sample(1:32, N, replace = TRUE)
  which_is_random <- runif(N) > S
  innovation_i[which_is_random] <- innovation_random[which_is_random]
```

The same must be done for `update_social_behaviours()`, and we need to add the parameter $S$ in the pertinent function calls. We can now re-source [main.R](main.R). 

```{r message=FALSE}
source("main.R")
```

And call the new function. We can first test $S=1$ (so this should produce the same results we already have) with the "relasitic" value of $\alpha_g$ and $\alpha_e$ we know produce a result in line with Whiten et al., 1999.

```{r}
test <- run_oranzees(t_max = 6000, alpha_e = 0.8, alpha_g = 0.2, S = 1, init_world = TRUE, n_run = 5)
out <- analyse_patterns(test)
out$n[16:20]
```

Let's see what happens with $S=0.5$ and $S=0$:

```{r}
test <- run_oranzees(t_max = 6000, alpha_e = 0.8, alpha_g = 0.2, S = 0.5, init_world = TRUE, n_run = 5)
out <- analyse_patterns(test)
out$n[16:20]

test <- run_oranzees(t_max = 6000, alpha_e = 0.8, alpha_g = 0.2, S = 0, init_world = TRUE, n_run = 5)
out <- analyse_patterns(test)
out$n[16:20]
```

As expected, the number of cultural behaviours decreased (apparently of a similar amount, with both values of $S$). We can decide whether we want to study the effect of this parameters in the final analysis.

***

