---
title: "Oranzees first tests results"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Codes for behavioural patterns

Whiten et al., 1999 extract the above patterns from their data:

* *A*: patterns absent at no site.

* *B*: patterns not achieving habitual frequencies at any site.

* *C*: patterns for which any absence can be explained by local ecological factors.

* *D*: patterns customary or habitual at some sites yet absent at others, with no ecological explanation, i.e. the "cultural" behaviours.

We thus need to write a further function, `analyse_table()`, in [test.R](test.R), that takes as input the output of `run_oranzees()`, and produces the proportion of "cultural" behaviours in each run:

```{r}
analyse_table <- function(my_test){
  output <- my_test %>%
    group_by(behaviour, run) %>%
    summarise(A = !("absent" %in% code) && !("ecological explanation" %in% code), 
              B = !("habitual" %in% code) && !("customary" %in% code),
              C = !("absent" %in% code) && ("ecological explanation" %in% code)) %>%
    ungroup() %>%
    mutate (D = !(A | B | C)) %>%
    group_by(run) %>%
    count(D) %>%
    filter(D == TRUE) 
  output <- output$n/38
}
```

### Running the tests

Before starting, we need all the functions present in [main.R](main.R), as well as the library `tictoc`, to have information on the time needed for the simulations to run.

```{r message=FALSE}
library(tictoc)
source("main.R")
```

We run 20 simulations for $t_\text{max}=12000$, reinitialising the world for each simulations, for the following combination of parameters:

$opt$ | $\alpha_g$ 
------|-------
0     | 0.5
0     | 0.6
0     | 0.7
0.1   | 0.5
0.1   | 0.6 
0.1   | 0.7

The code, for example for the first combination, is:

```{r eval=FALSE}
tic()
test_0.5_0 <- run_oranzees(t_max = 12000, opt = 0, alpha_g = 0.5, init_world = 1, n_run = 20)
toc()
write_csv(test_0.5_0, "output_test/test_0.5_0.csv")
```

We run seprately the simulations (see [test.R](test.R)), and simply load here the results:

```{r message=FALSE}
test_0_05 <-analyse_table(read_csv("output_test/test_0_0.5.csv"))
test_0_06 <-analyse_table(read_csv("output_test/test_0_0.6.csv"))
test_0_07 <-analyse_table(read_csv("output_test/test_0_0.7.csv"))
test_01_05 <-analyse_table(read_csv("output_test/test_0.1_0.5.csv"))
test_01_06 <-analyse_table(read_csv("output_test/test_0.1_0.6.csv"))
test_01_07 <-analyse_table(read_csv("output_test/test_0.1_0.7.csv"))
```

We can now plot the proportion of traits considered "cultural", according to the patterns in Whiten et al., 1999 in each condition:

```{r}
tibble(opt = rep(c("0", "0.1"), each = 60), 
       alpha_g = rep(c("0.5", "0.6", "0.7", "0.5", "0.6", "0.7"), each = 20),
      behaviours = c(test_0_05, test_0_06, test_0_07, test_01_05, test_01_06, test_01_07)) %>%
  ggplot(aes( x = opt, y = behaviours, fill = alpha_g)) +
    geom_boxplot(outlier.shape = NA) +
    geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1)) +
  scale_fill_manual(values = c("#FFDB6D", "#F4EDCA", "#D16103")) +
    theme_bw() +
    labs(y = "proportion of traits considered 'cultural'") 
```

Whiten et al. 1999 found 38 "cultural" behaviour out of a total of 65 considered, i.e. a proportion of around $0.58$. This is close to our conditions with no optimisation and absent or low genetic influence. It is useful to look more in details to these conditions to have a clearer picture of how the results look like. 

First, we can use the `plot_oranzees()` function to plot the distribution of behaviours in single runs:

```{r message = FALSE}
test_0_0.5 <- read_csv("output_test/test_0_0.5.csv")
```

```{r fig.height = 7}
plot_oranzees(filter(test_0_0.5, run == 1), social=TRUE)
```

```{r fig.height = 9}
plot_oranzees(filter(test_0_0.5, run == 1), social=FALSE)
```

***
### Code modification 1:

During the tests a bus was discovered and the code in [main.R](main.R) changed accordingly:

Added checks on adults, subadults, and juveniles subsets in `run_oranzees()` and `test_oranzees_2()`. With small populations, the subsets could be equal to 0, generating an error. As they need to be a "majority", I put a check at `>= 3` 

```{r eval=FALSE}
# customary:
customary_adults <- rep(FALSE, 38)
if(sum(adults) >= 3){
  customary_adults <- colSums(pop[adults,1:38])>(sum(adults)/2)
}
customary_subadults <- rep(FALSE, 38)
if(sum(subadults) >= 3){
  customary_subadults <- colSums(pop[subadults,1:38])>(sum(subadults)/2)
}
customary_juveniles <- rep(FALSE, 38)
if(sum(juveniles) >= 3){
  customary_juveniles <- colSums(pop[juveniles,1:38])>(sum(juveniles)/2)
}
```

***