---
title: "Oranzees first tests results"
author: "Alberto Acerbi"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


### Codes for behavioural patterns

Whiten et al., 1999 extract the above patterns from their data:

* *A*: patterns absent at no site.

* *B*: patterns not achieving habitual frequencies at any site.

* *C*: patterns for which any absence can be explained by local ecological factors.

* *D*: patterns customary or habitual at some sites yet absent at others, with no ecological explanation, i.e. the "cultural" behaviours.

We thus need to write a further function, `analyse_patterns()`, in [test.R](test.R), that takes as input the output of `run_oranzees()`, and produces the number of behaviours for each pattern in each run:

```{r}
analyse_patterns <- function(my_test){
  output <- tibble(pattern = rep(c("A", "B", "C", "D"), each = max(my_test$run)), 
                   n = rep(NA, max(my_test$run) * 4))
  process <- my_test %>%
    group_by(behaviour, run) %>%
    summarise(A = !("absent" %in% code) && !("ecological explanation" %in% code), 
              B = !("habitual" %in% code) && !("customary" %in% code),
              C = !("absent" %in% code) && ("ecological explanation" %in% code)) %>%
    ungroup() %>%
    mutate (D = !(A | B | C)) %>%
    group_by(run)
  
  temp <- count(process, A) %>%
    filter(A == TRUE)
  out <- temp$n
  if(length(temp$n) < max(my_test$run)){
    out <- c(temp$n, rep(0, max(my_test$run)-length(temp$n)))
  }
  output[output$pattern=="A",]$n <- out  
  
  temp <- count(process, B) %>%
    filter(B == TRUE)
  out <- temp$n
  if(length(temp$n) < max(my_test$run)){
    out <- c(temp$n, rep(0, max(my_test$run)-length(temp$n)))
  }
  output[output$pattern=="B",]$n <- out  
  
  temp <- count(process, C) %>%
    filter(C == TRUE)
  out <- temp$n
  if(length(temp$n) < max(my_test$run)){
    out <- c(temp$n, rep(0, max(my_test$run)-length(temp$n)))
  }
  output[output$pattern=="C",]$n <- out 
  
  temp <- count(process, D) %>%
    filter(D == TRUE)
  out <- temp$n
  if(length(temp$n) < max(my_test$run)){
    out <- c(temp$n, rep(0, max(my_test$run)-length(temp$n)))
  }
  output[output$pattern=="D",]$n <- out 
  
  output
}

```

### Running the tests

Before starting, we need all the functions present in [main.R](main.R), as well as the library `tictoc`, to have information on the time needed for the simulations to run.

```{r message=FALSE}
library(tictoc)
source("main.R")
```

We run 20 simulations for $t_\text{max}=6000$, reinitialising the world for each simulations, for the following combination of parameters:

$\alpha_e$ | $\alpha_g$ 
------|-------
0     | 0
0     | 0.5
0     | 1
0.5   | 0
0.5   | 0.5 
0.5   | 1
1     | 0
1     | 0.5
1     | 1

The code, for example for the first combination, is:

```{r eval=FALSE}
tic()
test <- run_oranzees(t_max = 6000, alpha_e = 0, alpha_g = 0, init_world = 1, n_run = 20)
toc()
write_csv(test, "output_test/test_0_0.csv")
```

We run seprately the simulations (see [test.R](test.R)), and simply load here the results:

```{r message=FALSE}
test_0_0 <-analyse_patterns(read_csv("output_test/test_0_0.csv"))
test_0_05 <-analyse_patterns(read_csv("output_test/test_0_05.csv"))
test_0_1 <-analyse_patterns(read_csv("output_test/test_0_1.csv"))
test_05_0 <-analyse_patterns(read_csv("output_test/test_05_0.csv"))
test_05_05 <-analyse_patterns(read_csv("output_test/test_05_05.csv"))
test_05_1 <-analyse_patterns(read_csv("output_test/test_05_1.csv"))
test_1_0 <-analyse_patterns(read_csv("output_test/test_1_0.csv"))
test_1_05 <-analyse_patterns(read_csv("output_test/test_1_05.csv"))
test_1_1 <-analyse_patterns(read_csv("output_test/test_1_1.csv"))
```

We can now plot the number of traits considered "cultural", according to the patterns in Whiten et al., 1999 in each condition:

```{r}
tibble(alpha_e = rep(c("0", "0.5", "1"), each = 60), 
       alpha_g = rep(c("0", "0.5", "1","0", "0.5", "1","0", "0.5", "1"), each = 20),
      behaviours = c(test_0_0[test_0_0$pattern=="D",]$n, 
                     test_0_05[test_0_05$pattern=="D",]$n,
                     test_0_1[test_0_1$pattern=="D",]$n,
                     test_05_0[test_05_0$pattern=="D",]$n,
                     test_05_05[test_05_05$pattern=="D",]$n,
                     test_05_1[test_05_1$pattern=="D",]$n,
                     test_1_0[test_1_0$pattern=="D",]$n,
                     test_1_05[test_1_05$pattern=="D",]$n,
                     test_1_1[test_1_1$pattern=="D",]$n)) %>%
  ggplot(aes( x = alpha_g, y = behaviours, fill = alpha_e)) +
    geom_boxplot(outlier.shape = NA) +
    geom_point(pch = 21, position = position_jitterdodge(jitter.width = 0.1)) +
  scale_fill_manual(values = c("#FFDB6D", "#F4EDCA", "#D16103")) +
    theme_bw() +
    labs(y = "Number of traits considered 'cultural'") 
```

Whiten et al., 1999 found 38 "cultural" behaviour out of a total of 65 considered. This is close, for these parameters, to the conditions with high ecological variability ($\alpha_e=1$) and no genetic differences ($\alpha_g=0$), or to the condition with no ecological variability ($\alpha_e=0$) and intermediate genetic differences ($\alpha_g=0.5$).  It is useful to look more in details to this conditions to have a clearer picture of how the results look like. 

For example, what about the other patterns? We can plot all the pattern proportions in the conditions we are interested, and compare them directly with the results of Whiten et al., 1999.

```{r}
whiten <- tibble(pattern = 1:4, 
                 n = c(7, 16, 3, 38), 
                 data = rep("Whiten", 4)) 
ggplot(data = test_1_0, aes(x = pattern, y = n)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width=0.05, alpha=0.5) +
    theme_bw() +
    geom_line(data = whiten, colour="red") +
    geom_point(data = whiten, colour="red") 
```

```{r}
whiten <- tibble(pattern = 1:4, 
                 n = c(7, 16, 3, 38), 
                 data = rep("Whiten", 4)) 
ggplot(data = test_0_05, aes(x = pattern, y = n)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width=0.05, alpha=0.5) +
    theme_bw() +
    geom_line(data = whiten, colour="red") +
    geom_point(data = whiten, colour="red") 
```
